---
title: "Plot different alphas"
author: "Joaquim Aguirre-Plans"
date: "4/16/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Plot different alphas of a given model.

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
require(magrittr)
library(tidyr)
library(igraph)
set.seed(1510)
`%ni%` <- Negate(`%in%`)
```

### Define variables

```{r}
input_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/scripts/SampleSizeShiny/data/example_pearson_pval_0.05'
plots_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots'
tables_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/tables'

dataset_selected = "gtex:whole.blood"
model_selected = "Stretched exponential (by linear fit)"
N_vals = seq(10, 50000, 10)

# Define palette using the scale_fill_brewer Oranges palette and selecting manually the colors using RColorBrewer
# https://stackoverflow.com/questions/29466239/how-to-set-the-color-range-of-scale-colour-brewer-in-ggplot2-palette-selecte 
blue_palette = RColorBrewer::brewer.pal(n = 9, "Blues") # I get 8 and exclude the 4 first
name2color <- c(
  "0" = "red",
  "0.5" = blue_palette[2],
  "1" = blue_palette[3],
  "1.25" = blue_palette[4],
  "1.5" = blue_palette[6],
  "1.65" = blue_palette[7],
  #"1.75" = blue_palette[7],
  "2" = blue_palette[9]
)
```

Read files:

```{r}
analytical_summary_file = paste(input_dir, 'analytical_model_summary_pearson_pval_0.05.txt', sep='/')
analytical_model_summary_df = fread(analytical_summary_file)
analytical_model_selected = analytical_model_summary_df %>% 
  dplyr::select(type_dataset, model, a, b, L, max_value_in_dataset, unnorm_L) %>% 
  filter((type_dataset == dataset_selected) & (model == model_selected)) %>% unique()
analytical_model_selected
```

### Run Test

```{r}
#'  calculate_predictions_using_stretched_exponential_model_optimized
#'  Formula to calculate exponential decay of significant interactions from a list of sample sizes.
#'  This formula requires the use of the L parameter
#'  @param x List of sample sizes.
#'  @param a Slope coefficient.
#'  @param b Intercept coefficient.
#'  
calculate_predictions_using_stretched_exponential_model_optimized = function(x, L, a, b){
  y = L * exp((b*x**(-a+1))/(-a+1))
  #y = exp(log(L) - exp(b+a*log(x)))
  return(y)
}

#'  calculate_prediction_from_analytical_model
#'  Function to calculate predictions using a specific analytical model
#'  @param model Name of the model used.
#'  @param x_list List of values in the x axis (e.g. size)
#'  @param a Parameter a.
#'  @param b Parameter b.
#'  @param L Parameter L.
#'  
calculate_prediction_from_analytical_model = function(model, x_list, a, b, L){
  if(model == "Logarithmic"){
    prediction_result = (log(x_list)*a + b)
  } else if(model == "Linear"){
    prediction_result = (x_list*a + b)
  } else if(model == "Exponential"){
    prediction_result = (exp((x_list*a + b)))
  } else if(model == "Square root"){
    prediction_result = (exp((log(x_list)*a + b)))
  } else if((model == "Stretched exponential (by optimization)") | (model == "Stretched exponential (by linear fit)")){
    prediction_result = calculate_predictions_using_stretched_exponential_model_optimized(x=x_list, L=L, a=a, b=b)
  } else if(model == "Stretched exponential (without L)"){
    prediction_result = calculate_predictions_using_stretched_exponential_model_without_L(x=x_list, a=a, b=b)
  }
  return(prediction_result)
}
```

For different alphas and sample sizes, calculate the predictions of the analytical model:

```{r}
list_alphas = c(#1, 
                1.25, 
                1.5, 
                round(analytical_model_selected$a, 2),
                2)
cols = c("model", "model_result", "size", "a", "L", "max_value_in_dataset")
predicted_results_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))
for (alpha_selected in list_alphas){
  prediction_result = 
    calculate_prediction_from_analytical_model(model=model_selected, 
                                               x_list=N_vals, 
                                               a=alpha_selected, 
                                               b=analytical_model_selected$b, 
                                               L=analytical_model_selected$L)
  prediction_result = prediction_result * analytical_model_selected$max_value_in_dataset
  predicted_results_df = rbind(predicted_results_df, data.frame(model=model_selected, model_result=prediction_result, size=N_vals, a=alpha_selected, L=analytical_model_selected$L, max_value_in_dataset=analytical_model_selected$max_value_in_dataset))
}

predicted_results_df_file = paste(tables_dir, 
                                "/comparison_alphas_",
                                dataset_selected,
                                ".txt",
                                sep = "")

predicted_results_df %>% 
  fwrite(predicted_results_df_file)

head(predicted_results_df)
```

Plot the results comparing the analytical model with different alphas:

```{r}
N_vals = seq(10, 5000, 10)

predicted_results_df = predicted_results_df %>%
  filter(size %in% N_vals) %>%
  group_by(a) %>% mutate(norm = (model_result / max_value_in_dataset)/L) %>% ungroup() %>%
  mutate(a = factor(a, levels = as.character(sort(unique(predicted_results_df$a)))))

comparison_alphas_plot = predicted_results_df %>%
  ggplot(aes(x=log10(size), y=norm, col=a)) +
  geom_line(size=1) +
  scale_color_manual(name = expression(alpha), 
                     values = as.vector(name2color[levels(factor(predicted_results_df$a))])) +
  theme_linedraw() +
  xlab("Num. samples (Log10)") +
  ylab("Frac. significant correlations") +
  theme(aspect.ratio = 1, 
        plot.title =  element_text(size = 20, face="bold"), 
        axis.title = element_text(size = 17, face="bold"), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        legend.title=element_text(size=16, face="bold"),
        panel.grid.major=element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(family = "Helvetica")
        )

plot_file = paste(plots_dir,
                  "/comparison_alphas_",
                  dataset_selected,
                  ".png",
                  sep = "")
ggsave(
  plot_file,
  plot = comparison_alphas_plot,
  dpi = 1200,
  #width = 10000,
  height = 6000,
  units = c("px")
)
print(comparison_alphas_plot)
```

### Test using values from random distribution

We select random values from a dataset:

```{r}
random_dataset_file <- paste(tables_dir, 
                             "/random_dataset_correlations_Whole.Blood.txt",
                             sep = "")
network_file <- "/scratch/j.aguirreplans/Scipher/SampleSizeProject/networks_gtex/reads/Whole.Blood/pearson_gtex_Whole.Blood_size_100_rep_1.net"

if (!(file.exists(random_dataset_file))) {
  network_df <- fread(network_file)
  random_dataset_df <- network_df %>% slice_sample(n=1000000, replace = FALSE)
  random_dataset_df %>% fwrite(random_dataset_file)
} else {
  random_dataset_df <- fread(random_dataset_file)
}

```

```{r}
blue_palette = RColorBrewer::brewer.pal(n = 9, "Blues") # I get 8 and exclude the 4 first
name2color <- c(
  "0" = "red",
  "0.5" = blue_palette[2],
  "1" = blue_palette[3],
  "1.25" = blue_palette[4],
  "1.5" = blue_palette[6],
  "1.65" = blue_palette[7],
  #"1.75" = blue_palette[7],
  "2" = blue_palette[9]
)
get_frac_links <- function(prob_val, corr_distr) {
  return(length(corr_distr[corr_distr > prob_val]) / length(corr_distr))
}
list_alphas = c(1, 1.5, 2)
N_vals = seq(10, 1000, 10)
b = analytical_model_selected$b
#L = analytical_model_selected$unnorm_L
corr_distr = abs(random_dataset_df$score)
L = length(corr_distr)
cols = c("model", "model_result", "size", "a", "L")
predicted_results_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))

for (alpha_selected in list_alphas){
  prediction_prob = b * N_vals ** (-alpha_selected)
  prediction_result = unlist(lapply(X=prediction_prob, FUN=get_frac_links, corr_distr=corr_distr))
  predicted_results_df = rbind(predicted_results_df, data.frame(model=model_selected, model_result=prediction_result, size=N_vals, a=alpha_selected, L=L))
}

predicted_results_df = predicted_results_df %>%
  mutate(a = factor(a, levels = as.character(sort(unique(predicted_results_df$a)))))

comparison_alphas_plot = predicted_results_df %>%
  ggplot(aes(x=log10(size), y=model_result, col=a)) +
  geom_line(size=1) +
  scale_color_manual(name = expression(alpha), 
                     values = as.vector(name2color[levels(factor(predicted_results_df$a))])) +
  theme_linedraw() +
  xlab("Num. samples (Log10)") +
  ylab("Frac. significant correlations") +
  #scale_x_continuous(trans='log10') +
  theme(aspect.ratio = 1, 
        plot.title =  element_text(size = 20, face="bold"), 
        axis.title = element_text(size = 17, face="bold"), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        legend.title=element_text(size=16, face="bold"),
        panel.grid.major=element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(family = "Helvetica")
  )

plot_file = paste(plots_dir,
                  "/comparison_alphas_random_dataset_",
                  dataset_selected,
                  ".png",
                  sep = "")
ggsave(
  plot_file,
  plot = comparison_alphas_plot,
  dpi = 1200,
  #width = 10000,
  height = 6000,
  units = c("px")
)
print(comparison_alphas_plot)
```

Using different links for different sample sizes:

```{r}
random_dataset_file <- paste(tables_dir, 
                             "/random_dataset_correlations_different_sizes_Whole.Blood.txt",
                             sep = "")
if (!(file.exists(random_dataset_file))) {
  network_file <- "/scratch/j.aguirreplans/Scipher/SampleSizeProject/networks_gtex/reads/gtex_Whole.Blood_pearson_combined.txt"
  network_df <- fread(network_file, header = T)
  random_dataset_df <- network_df %>% slice_sample(n=1000000, replace = FALSE)
  rm(network_df)
  random_dataset_df %>% fwrite(random_dataset_file)
} else {
  random_dataset_df <- fread(random_dataset_file)
  cols <- colnames(random_dataset_df)[3:ncol(random_dataset_df)]
  split_strings <- strsplit(cols, "\\.")
  cols <- sapply(split_strings, function(x) as.numeric(x[1]))
  colnames(random_dataset_df) <- c("Node.1", "Node.2", as.character(cols))
}

```

```{r}
blue_palette = RColorBrewer::brewer.pal(n = 9, "Blues") # I get 8 and exclude the 4 first
name2color <- c(
  "0" = "red",
  "0.5" = blue_palette[2],
  "1" = blue_palette[3],
  "1.25" = blue_palette[4],
  "1.5" = blue_palette[6],
  "1.65" = blue_palette[7],
  #"1.75" = blue_palette[7],
  "2" = blue_palette[9]
)
list_alphas = c(1, 1.5, 2)
N_vals = as.numeric(colnames(random_dataset_df)[3:ncol(random_dataset_df)])
b = analytical_model_selected$b
#L = analytical_model_selected$unnorm_L
L = nrow(random_dataset_df)
cols = c("model_result", "size", "a", "L")
predicted_results_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))

for (alpha_selected in list_alphas){
  prediction_prob = b * N_vals ** (-alpha_selected)
  for (i in seq_along(N_vals)) {
    N <- N_vals[i]
    prob_val <- prediction_prob[i]
    corr_distr <- random_dataset_df %>% dplyr::select(as.character(N)) %>% dplyr::rename(prob = as.character(N))
    prediction_result <- nrow((corr_distr %>% filter(prob > prob_val))) / nrow(corr_distr)
    predicted_results_df = rbind(predicted_results_df, data.frame(model_result=prediction_result, size=N, a=alpha_selected, L=L))
  }
}

predicted_results_df = predicted_results_df %>%
  mutate(a = factor(a, levels = as.character(sort(unique(predicted_results_df$a)))))

comparison_alphas_plot = predicted_results_df %>%
  ggplot(aes(x=log10(size), y=model_result, col=a)) +
  geom_line(size=1) +
  scale_color_manual(name = expression(alpha), 
                     values = as.vector(name2color[levels(factor(predicted_results_df$a))])) +
  theme_linedraw() +
  xlab("Num. samples (Log10)") +
  ylab("Frac. significant correlations") +
  #scale_x_continuous(trans='log10') +
  theme(aspect.ratio = 1, 
        plot.title =  element_text(size = 20, face="bold"), 
        axis.title = element_text(size = 17, face="bold"), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        legend.title=element_text(size=16, face="bold"),
        panel.grid.major=element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(family = "Helvetica")
  )

plot_file = paste(plots_dir,
                  "/comparison_alphas_random_dataset_",
                  dataset_selected,
                  ".png",
                  sep = "")
ggsave(
  plot_file,
  plot = comparison_alphas_plot,
  dpi = 1200,
  #width = 10000,
  height = 6000,
  units = c("px")
)
print(comparison_alphas_plot)
```
