---
title: "Scipher co-expression networks analysis"
author: "Joaquim Aguirre-Plans"
date: "1/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Analyze the co-expression networks created from the Scipher's gene expression dataset.

```{r, message=FALSE, echo=FALSE, results='hide'}
library(data.table)
library(dplyr)
library(igraph)
library(ggplot2)
require(magrittr)
library(wTO)
set.seed(1510)
options(bitmapType='cairo')
`%ni%` <- Negate(`%in%`)
```

```{r define_files, echo=FALSE, results='hide'}
# Define working directories
databases_dir = '/home/j.aguirreplans/Databases'
ppi_dir = '/home/j.aguirreplans/data/PPI'
data_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/Dec2021'
networks_dir = '/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher'

# Define input files
#metadata_RNA_file = paste(data_dir, '00_data/scipher_metadata_rnaseq_counts.csv', sep='/')
#metadata_RNA_0m_file = paste(data_dir, '00_data/scipher_metadata_rnaseq_counts_0m.csv', sep='/')
ppi_network_symbol_file = paste(ppi_dir, 'interactome_2019_merged_symbols.csv', sep='/')
ppi_df = fread(ppi_network_symbol_file) %>% select(proteinA_symbol, proteinB_symbol)
ppi_net = graph_from_data_frame(ppi_df, directed=F) %>% simplify()

```

## Get correlation distributions

```{r test echo=FALSE, eval=FALSE}
scipher_coexpression_network_test_file = "/scratch/j.aguirreplans/spearman_coexpression_network_test.net"
scipher_histogram_plot_test_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_spearman_correlation_scipher_test.png"
scipher_histogram_cumulative_plot_test_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_spearman_correlation_scipher_test.png"
tmp_df = as.data.frame(fread(scipher_coexpression_network_test_file)) %>% dplyr::select(Node.1, Node.2, spearman, pval.adj)
tmp_df$scipher_webpower_cut = ifelse(abs(tmp_df$spearman) >= 0.32, "Corr. \u2265 0.32 (\u2264 565 samples)", "Corr. < 0.32 (> 565 samples)")
tmp_df$breaks =  cut(abs(tmp_df$spearman), breaks = seq(0,1,0.1), labels=seq(0,0.9,0.1))

print(length((tmp_df %>% filter(spearman >= 0.32))$spearman) / length(tmp_df$spearman)*100)
print(length((tmp_df %>% filter(pval.adj < 0.05))$spearman) / length(tmp_df$spearman)*100)

tmp_count_df = tmp_df %>% count(breaks, scipher_webpower_cut)
rm(tmp_df)
tmp_count_df$percent =  tmp_count_df$n / sum(tmp_count_df$n)
tmp_count_df$cumulative_percent =  rev(cumsum(rev(tmp_count_df$n / sum(tmp_count_df$n))))

tmp_df %>% 
  ggplot() + 
  geom_histogram(aes(x=spearman, y = stat(width*density)), binwidth=0.1, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Distribution of Spearman correlation in Scipher dataset") +
  xlab("Absolute Spearman correlation") + 
  ylab("% of edges")

tmp_count_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks, weight=percent), fill="#69b3a2") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Distribution of Spearman correlation in Scipher dataset") +
  xlab("Absolute Spearman correlation") + 
  ylab("% of edges")

ggsave(
  scipher_histogram_plot_test_file,
  dpi = 1200
)

tmp_count_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks, weight=cumulative_percent, fill=scipher_webpower_cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Spearman correlation in Scipher dataset") +
  xlab("Absolute Spearman correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title=""))

ggsave(
  scipher_histogram_cumulative_plot_test_file,
  dpi = 1200
)


```


```{r pearson_correlation_distribution_scipher, echo=FALSE, eval=FALSE}
scipher_coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/pearson_scipher_complete_dataset_all_samples.net"
scipher_histogram_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_pearson_correlation_scipher.png"
scipher_webpower_histogram_cumulative_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_pearson_correlation_scipher_webpower.png"
scipher_hiclimr_histogram_cumulative_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_pearson_correlation_scipher_hiclimr.png"
scipher_histogram_cumulative_plot_pval_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_pearson_correlation_by_pval_scipher.png"

coexpression_network_df = as.data.frame(fread(scipher_coexpression_network_file)) %>% dplyr::select(Node.1, Node.2, score, pval.adj)
coexpression_network_df$scipher_webpower_cut = ifelse(abs(coexpression_network_df$score) >= 0.32, "Corr. \u2265 0.32", "Corr. < 0.32")
coexpression_network_df$scipher_hiclimr_cut = ifelse(abs(coexpression_network_df$score) >= 0.081, "Corr. \u2265 0.081", "Corr. < 0.081")
#coexpression_network_df$scipher_webpower_cut = ifelse(abs(coexpression_network_df$score) >= 0.32, "Corr. \u2265 0.32 (\u2264 565 samples)", "Corr. < 0.32 (> 565 samples)")
coexpression_network_df$pval.adj.cut = ifelse(abs(coexpression_network_df$pval.adj) < 0.05, "P-value < 0.05", "P-value \u2265 0.05")
coexpression_network_df$breaks1 =  cut(coexpression_network_df$score, breaks = seq(-0.8,1,0.2), labels=seq(-0.8,0.8,0.2), include.lowest=TRUE)
coexpression_network_df$breaks2 =  cut(abs(coexpression_network_df$score), breaks = seq(0,1,0.1), labels=seq(0,0.9,0.1), include.lowest=TRUE)

print(length((coexpression_network_df %>% filter(score >= 0.32))$score) / length(coexpression_network_df$score)*100)
print(length((coexpression_network_df %>% filter(score >= 0.081))$score) / length(coexpression_network_df$score)*100)
print(length((coexpression_network_df %>% filter(pval.adj < 0.05))$score) / length(coexpression_network_df$score)*100)
print(max(abs((coexpression_network_df %>% filter(pval.adj >= 0.05))$score)))
print(min(abs((coexpression_network_df %>% filter(pval.adj < 0.05))$score)))

coexpression_network_count_df = coexpression_network_df %>% count(breaks1)
coexpression_network_count_abs_webpower_df = coexpression_network_df %>% count(breaks2, scipher_webpower_cut)
coexpression_network_count_abs_hiclimr_df = coexpression_network_df %>% count(breaks2, scipher_hiclimr_cut)
coexpression_network_count_abs_pval_df = coexpression_network_df %>% count(breaks2, pval.adj.cut)
#rm(coexpression_network_df)
coexpression_network_count_df$percent =  coexpression_network_count_df$n / sum(coexpression_network_count_df$n)
coexpression_network_count_abs_webpower_df$cumulative_percent =  rev(cumsum(rev(coexpression_network_count_abs_webpower_df$n / sum(coexpression_network_count_abs_webpower_df$n))))
coexpression_network_count_abs_hiclimr_df$cumulative_percent =  rev(cumsum(rev(coexpression_network_count_abs_hiclimr_df$n / sum(coexpression_network_count_abs_hiclimr_df$n))))
coexpression_network_count_abs_pval_df$percent =  coexpression_network_count_abs_pval_df$n / sum(coexpression_network_count_abs_pval_df$n)
coexpression_network_count_abs_pval_df = rbind((coexpression_network_count_abs_pval_df %>% filter(!(pval.adj.cut == "P-value < 0.05"))), (coexpression_network_count_abs_pval_df %>% filter(pval.adj.cut == "P-value < 0.05")))
coexpression_network_count_abs_pval_df$cumulative_percent =  rev(cumsum(rev(coexpression_network_count_abs_pval_df$n / sum(coexpression_network_count_abs_pval_df$n))))
coexpression_network_count_abs_pval_df$pval.adj.cut <- factor(coexpression_network_count_abs_pval_df$pval.adj.cut, levels=c("P-value \u2265 0.05", "P-value < 0.05"))

#coexpression_network_count_df %>% 
#  ggplot() + 
#  geom_bar(aes(x=breaks1, weight=percent), fill="#69b3a2") +
#  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
#  ggtitle("Distribution of Pearson correlation") +
#  xlab("Pearson correlation") + 
#  ylab("% of edges") + 
#  theme(legend.position = "none", plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), #axis.text = element_text(size = 15))

coexpression_network_df %>% 
  ggplot() + 
  geom_histogram(aes(x=score, y = stat(width*density)), binwidth=0.1, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Distribution of Pearson correlation") +
  xlab("Pearson correlation") + 
  ylab("% of edges") + 
  theme(legend.position = "none", plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15))

ggsave(
  scipher_histogram_plot_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

coexpression_network_count_abs_webpower_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks2, weight=cumulative_percent, fill=scipher_webpower_cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Pearson correlation") +
  xlab("Absolute Pearson correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title="")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

ggsave(
  scipher_webpower_histogram_cumulative_plot_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

coexpression_network_count_abs_hiclimr_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks2, weight=cumulative_percent, fill=scipher_hiclimr_cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Pearson correlation") +
  xlab("Absolute Pearson correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title="")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

ggsave(
  scipher_hiclimr_histogram_cumulative_plot_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

coexpression_network_count_abs_pval_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks2, weight=cumulative_percent, fill=pval.adj.cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Pearson correlation") +
  xlab("Absolute Pearson correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title="")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

ggsave(
  scipher_histogram_cumulative_plot_pval_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

```












```{r pearson_correlation_distribution_scipher_bonferroni, echo=FALSE, eval=FALSE}
scipher_coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/pearson_scipher_complete_dataset_all_samples.net"
scipher_histogram_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_pearson_correlation_bonferroni_scipher.png"
scipher_webpower_histogram_cumulative_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_pearson_correlation_bonferroni_scipher_webpower.png"
scipher_hiclimr_histogram_cumulative_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_pearson_correlation_bonferroni_scipher_hiclimr.png"
scipher_histogram_cumulative_plot_pval_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_cumulative_pearson_correlation_bonferroni_by_pval_scipher.png"

coexpression_network_df = as.data.frame(fread(scipher_coexpression_network_file)) %>% dplyr::select(score, pvalue, pval.adj)
coexpression_network_df$pval.adj.bonf = pval.adjust(coexpression_network_df$pvalue, method = "bonferroni")
coexpression_network_df$scipher_webpower_cut = ifelse(abs(coexpression_network_df$score) >= 0.32, "Corr. \u2265 0.32", "Corr. < 0.32")
coexpression_network_df$scipher_hiclimr_cut = ifelse(abs(coexpression_network_df$score) >= 0.081, "Corr. \u2265 0.081", "Corr. < 0.081")
#coexpression_network_df$scipher_webpower_cut = ifelse(abs(coexpression_network_df$score) >= 0.32, "Corr. \u2265 0.32 (\u2264 565 samples)", "Corr. < 0.32 (> 565 samples)")
coexpression_network_df$pval.adj.cut = ifelse(abs(coexpression_network_df$pval.adj.bonf) < 0.05, "P-value < 0.05", "P-value \u2265 0.05")
coexpression_network_df$breaks1 =  cut(coexpression_network_df$score, breaks = seq(-0.8,1,0.2), labels=seq(-0.8,0.8,0.2), include.lowest=TRUE)
coexpression_network_df$breaks2 =  cut(abs(coexpression_network_df$score), breaks = seq(0,1,0.1), labels=seq(0,0.9,0.1), include.lowest=TRUE)

print(length((coexpression_network_df %>% filter(score >= 0.32))$score) / length(coexpression_network_df$score)*100)
print(length((coexpression_network_df %>% filter(score >= 0.081))$score) / length(coexpression_network_df$score)*100)
print(length((coexpression_network_df %>% filter(pval.adj.bonf < 0.05))$score) / length(coexpression_network_df$score)*100)
print(max(abs((coexpression_network_df %>% filter(pval.adj.bonf >= 0.05))$score)))
print(min(abs((coexpression_network_df %>% filter(pval.adj.bonf < 0.05))$score)))

coexpression_network_count_df = coexpression_network_df %>% count(breaks1)
coexpression_network_count_abs_webpower_df = coexpression_network_df %>% count(breaks2, scipher_webpower_cut)
coexpression_network_count_abs_hiclimr_df = coexpression_network_df %>% count(breaks2, scipher_hiclimr_cut)
coexpression_network_count_abs_pval_df = coexpression_network_df %>% count(breaks2, pval.adj.cut)
#rm(coexpression_network_df)
coexpression_network_count_df$percent =  coexpression_network_count_df$n / sum(coexpression_network_count_df$n)
coexpression_network_count_abs_webpower_df$cumulative_percent =  rev(cumsum(rev(coexpression_network_count_abs_webpower_df$n / sum(coexpression_network_count_abs_webpower_df$n))))
coexpression_network_count_abs_hiclimr_df$cumulative_percent =  rev(cumsum(rev(coexpression_network_count_abs_hiclimr_df$n / sum(coexpression_network_count_abs_hiclimr_df$n))))
coexpression_network_count_abs_pval_df$percent =  coexpression_network_count_abs_pval_df$n / sum(coexpression_network_count_abs_pval_df$n)
coexpression_network_count_abs_pval_df = rbind((coexpression_network_count_abs_pval_df %>% filter(!(pval.adj.cut == "P-value < 0.05"))), (coexpression_network_count_abs_pval_df %>% filter(pval.adj.cut == "P-value < 0.05")))
coexpression_network_count_abs_pval_df$cumulative_percent =  rev(cumsum(rev(coexpression_network_count_abs_pval_df$n / sum(coexpression_network_count_abs_pval_df$n))))
coexpression_network_count_abs_pval_df$pval.adj.cut <- factor(coexpression_network_count_abs_pval_df$pval.adj.cut, levels=c("P-value \u2265 0.05", "P-value < 0.05"))

#coexpression_network_count_df %>% 
#  ggplot() + 
#  geom_bar(aes(x=breaks1, weight=percent), fill="#69b3a2") +
#  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
#  ggtitle("Distribution of Pearson correlation") +
#  xlab("Pearson correlation") + 
#  ylab("% of edges") + 
#  theme(legend.position = "none", plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), #axis.text = element_text(size = 15))

coexpression_network_df %>% 
  ggplot() + 
  geom_histogram(aes(x=score, y = stat(width*density)), binwidth=0.1, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Distribution of Pearson correlation") +
  xlab("Pearson correlation") + 
  ylab("% of edges") + 
  theme(legend.position = "none", plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15))

ggsave(
  scipher_histogram_plot_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

coexpression_network_count_abs_webpower_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks2, weight=cumulative_percent, fill=scipher_webpower_cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Pearson correlation") +
  xlab("Absolute Pearson correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title="")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

ggsave(
  scipher_webpower_histogram_cumulative_plot_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

coexpression_network_count_abs_hiclimr_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks2, weight=cumulative_percent, fill=scipher_hiclimr_cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Pearson correlation") +
  xlab("Absolute Pearson correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title="")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

ggsave(
  scipher_hiclimr_histogram_cumulative_plot_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

coexpression_network_count_abs_pval_df %>% 
  ggplot() + 
  geom_bar(aes(x=breaks2, weight=cumulative_percent, fill=pval.adj.cut), alpha=0.9, color="#e9ecef", position = 'identity') + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L)) +
  ggtitle("Cumulative distribution of Pearson correlation") +
  xlab("Absolute Pearson correlation") + 
  ylab("Cumulative % of edges") +
  guides(fill=guide_legend(title="")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

ggsave(
  scipher_histogram_cumulative_plot_pval_file,
  dpi = 1200,
  width = 8750,
  height = 6000,
  units = c("px")
)

```











```{r comparison_of_correlation_distributions}
# Parse correlation of networks from different number of samples
num_samples_list = c(20, 40, 100, 200, 500)
networks_dir = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset"
#network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/pearson_scipher_complete_dataset_size_20_rep_1.net"
network_file = paste(networks_dir, paste("pearson_scipher_complete_dataset_size_", num_samples_list[1], "_rep_1.net", sep=""), sep="/")
pval_df = data.frame(matrix(ncol=3,nrow=0, dimnames=list(NULL, c("num_samples", "min_corr_sig", "max_corr_not_sig"))))
coexpression_df = as.data.frame(fread(network_file)) %>% dplyr::select(Node.1, Node.2, score, pval.adj)
pval_df = rbind(pval_df, data.frame(num_samples=num_samples_list[1], min_corr_sig=min(abs((coexpression_df %>% filter(pval.adj < 0.05))$score)), max_corr_not_sig=max(abs((coexpression_df %>% filter(pval.adj >= 0.05))$score))))
#coexpression_df %<>% dplyr::select(Node.1, Node.2, score) %>% dplyr::rename(!!as.character(num_samples_list[1]) := "score")
coexpression_df %<>% dplyr::select(Node.1, Node.2, score) %>% mutate(num_samples=num_samples_list[1])

for(num_samples in num_samples_list){
  
  if(!(as.character(num_samples) %in% names(coexpression_df))){
    network_file = paste(networks_dir, paste("pearson_scipher_complete_dataset_size_", num_samples, "_rep_1.net", sep=""), sep="/")
    selected_coexpression_df = as.data.frame(fread(network_file)) %>% dplyr::select(Node.1, Node.2, score, pval.adj)
    pval_df = rbind(pval_df, data.frame(num_samples=num_samples, min_corr_sig=min(abs((selected_coexpression_df %>% filter(pval.adj < 0.05))$score)), max_corr_not_sig=max(abs((selected_coexpression_df %>% filter(pval.adj >= 0.05))$score))))
    #selected_coexpression_df %<>% dplyr::select(Node.1, Node.2, score) %>% dplyr::rename(!!as.character(num_samples) := "score")
    #coexpression_df = cbind(coexpression_df, (selected_coexpression_df %>% dplyr::select(!!as.character(num_samples))))
    selected_coexpression_df %<>% dplyr::select(Node.1, Node.2, score) %>% mutate(num_samples=num_samples)
    coexpression_df = rbind(coexpression_df, selected_coexpression_df)
    rm(selected_coexpression_df)
  }

}
print(pval_df)

# Plot histogram
coexpression_df$num_samples = as.character(coexpression_df$num_samples)
coexpression_df$num_samples <- factor(coexpression_df$num_samples , levels=as.character(sort(as.integer(unique(coexpression_df$num_samples)))))

histogram_correlation_distribution = coexpression_df %>%
  ggplot(aes(score, color=num_samples)) +
  geom_freqpoly(binwidth=0.1) +
  scale_y_log10() + 
  ggtitle("Distribution of Pearson correlation") +
  labs(x="Pearson correlation", y = "Log (Number of edges)", color = "Number of samples") +
  theme_linedraw() + 
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")

histogram_correlation_distribution_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/distribution_log_pearson_scipher_complete_dataset.png"
ggsave(
  filename=histogram_correlation_distribution_file,
  plot=histogram_correlation_distribution,
  dpi = 1200
)

histogram_correlation_distribution 

```



```{r, eval=FALSE}

test_coexpression_df = rbind(((coexpression_df %>% filter(num_samples == 20))[1:100,]), ((coexpression_df %>% filter(num_samples == 120))[1:100,]), ((coexpression_df %>% filter(num_samples == 220))[1:100,]), ((coexpression_df %>% filter(num_samples == 520))[1:100,]))
test_coexpression_df$num_samples = as.character(test_coexpression_df$num_samples)
test_coexpression_df$num_samples <- factor(test_coexpression_df$num_samples , levels=c(as.character(sort(as.integer(unique(test_coexpression_df$num_samples[!(test_coexpression_df$num_samples == "all")])))), "all"))

test_coexpression_df %>%
  ggplot(aes(score, color=num_samples)) +
  geom_freqpoly(binwidth=0.1) +
  ggtitle("Distribution of Pearson correlation") +
  labs(x="Pearson correlation", y = "Number of edges", color = "Number of samples") +
  theme_linedraw() + 
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.position="bottom")


```






```{r pearson_correlation_distribution_gtex, echo=FALSE, eval=FALSE}
gtex_coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/pearson_scipher_complete_dataset_all_samples.net"
gtex_correlation_plot_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots/histogram_pearson_correlation_gtex_whole_blood.png"

gtex_coexpression_network_df = as.data.frame(fread(gtex_coexpression_network_file))

gtex_coexpression_network_df %>% 
  ggplot(aes(x=score)) + 
  geom_histogram( binwidth=0.1, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle("Distribution of Pearson correlation in GTEx Whole Blood dataset (742 samples)") +
  xlab("Pearson correlation") + 
  ylab("Counts")

ggsave(
  gtex_correlation_plot_file,
  dpi = 1200
)

rm(gtex_coexpression_network_df)

```

```{r read_coexpression_networks, echo=FALSE, results='hide', eval=FALSE}
# Read gene co-expression networks
read_coexpression_network <- function(coexpression_network_file){
  coexpression_network_df = as.data.frame(fread(coexpression_network_file))
  if (!(grepl('wto', coexpression_network_file, fixed = TRUE))){
    # Pass the matrix to the wTO in-line format
    rownames(coexpression_network_df) = colnames(coexpression_network_df)
    coexpression_network_df <- coexpression_network_df[order(rownames(coexpression_network_df)), order(colnames(coexpression_network_df))]
    coexpression_network_df = coexpression_network_df %>% wTO.in.line() %>% rename(score=wTO)
  } else {
    coexpression_network_df = coexpression_network_df %>% rename(score=wTO)
    coexpression_network_df = coexpression_network_df %>% filter(pval.adj < 0.05)
  }
  return(coexpression_network_df)
}

```

```{r, eval=FALSE}
coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/wgcna_scipher_complete_dataset_size_10_rep_1.net"
coexpression_network_df = read_coexpression_network(coexpression_network_file)

coexpression_network_df %>% 
  ggplot(aes(x=score)) + 
  geom_histogram( binwidth=0.05, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle("Distribution of co-expression (Sample size 10)") +
  xlab("Co-expression score") + 
  ylab("Counts")

```

Number of edges of the co-expression network of sample size 10 when we use the following thresholds:

<>* 0.5: `r #nrow(coexpression_network_df %>% filter(score > 0.5))`
<>* 0.6: `r #nrow(coexpression_network_df %>% filter(score > 0.6))`
<>* 0.7: `r #nrow(coexpression_network_df %>% filter(score > 0.7))`

```{r, eval=FALSE}
coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/wgcna_scipher_complete_dataset_size_100_rep_1.net"
rm(coexpression_network_df)
coexpression_network_df = read_coexpression_network(coexpression_network_file)

coexpression_network_df %>% 
  ggplot(aes(x=score)) + 
  geom_histogram( binwidth=0.05, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle("Distribution of co-expression (Sample size 100)") +
  xlab("Co-expression score") + 
  ylab("Counts")

```

Number of edges of the co-expression network of sample size 100 when we use the following thresholds:

* 0.5: `r #nrow(coexpression_network_df %>% filter(score > 0.5))`
* 0.6: `r #nrow(coexpression_network_df %>% filter(score > 0.6))`
* 0.7: `r #nrow(coexpression_network_df %>% filter(score > 0.7))`

```{r, eval=FALSE}
coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/wgcna_scipher_complete_dataset_size_300_rep_1.net"
rm(coexpression_network_df)
coexpression_network_df = read_coexpression_network(coexpression_network_file)

coexpression_network_df %>% 
  ggplot(aes(x=score)) + 
  geom_histogram( binwidth=0.05, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle("Distribution of co-expression (Sample size 300)") +
  xlab("Co-expression score") + 
  ylab("Counts")

```

Number of edges of the co-expression network of sample size 300 when we use the following thresholds:

* 0.5: `r #nrow(coexpression_network_df %>% filter(score > 0.5))`
* 0.6: `r #nrow(coexpression_network_df %>% filter(score > 0.6))`
* 0.7: `r #nrow(coexpression_network_df %>% filter(score > 0.7))`


```{r, eval=FALSE}
coexpression_network_file = "/scratch/j.aguirreplans/Scipher/SampleSize/networks_scipher/complete_dataset/wgcna_scipher_complete_dataset_size_500_rep_1.net"
rm(coexpression_network_df)
coexpression_network_df = read_coexpression_network(coexpression_network_file)

coexpression_network_df %>% 
  ggplot(aes(x=score)) + 
  geom_histogram( binwidth=0.05, fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle("Distribution of co-expression (Sample size 500)") +
  xlab("Co-expression score") + 
  ylab("Counts")

```

Number of edges of the co-expression network of sample size 500 when we use the following thresholds:

* 0.5: `r #nrow(coexpression_network_df %>% filter(score > 0.5))`
* 0.6: `r #nrow(coexpression_network_df %>% filter(score > 0.6))`
* 0.7: `r #nrow(coexpression_network_df %>% filter(score > 0.7))`

