---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(data.table)
library(dplyr)
library(gghalves)
library(ggplot2)
library(igraph)
require(magrittr)
library(stringr)
library(tidyr)
set.seed(1510)
`%ni%` <- Negate(`%in%`)
```

```{r}
dataset = "gtex"
type_dataset = "Whole.Blood"
sample_group_type = "tumor" # if dataset = "tcga"
type_counts = "reads" # tpm, reads
sizes_selected = c(20, 100, 200, 300)
threshold_selected = 0.05
```

### Read the co-expression networks across different sample sizes and extract the co-expression

Fetch the files containing the gene co-expression networks of different sample sizes:

```{r}
coexpression_dir = sprintf("/scratch/j.aguirreplans/Scipher/SampleSize/networks_%s/reads/%s/consensus", dataset, type_dataset)
all_coexpression_files = list.files(coexpression_dir)
coexpression_files_selected = data.frame(matrix(ncol=length(c("size", "file")),nrow=0, dimnames=list(NULL, c("size", "file"))))

for (coexpression_file in all_coexpression_files){
  file_split = strsplit(gsub(".net", "", coexpression_file), split="_")[[1]]
  size = as.numeric(file_split[(length(file_split)-1)])
  if (size %in% sizes_selected){
    coexpression_files_selected = rbind(coexpression_files_selected, data.frame(size=size, file=paste(coexpression_dir, coexpression_file, sep="/")))
  }
}
```

Read the co-expression networks and extract the co-expression.
Keep only co-expression values with significant p-value (above 0.05). If not, the value is set to 0.

```{r}
coexpression_file_merged = sprintf("/scratch/j.aguirreplans/Scipher/SampleSize/networks_gtex/reads/%s_%s_pearson_consensus_combined_reduced.txt", dataset, type_dataset)
if (!(file.exists(coexpression_file))){
  for (i in 1:length(sort(coexpression_files_selected$size))){
    size_selected = sort(coexpression_files_selected$size)[i]
    coexpression_file = (coexpression_files_selected %>% filter(size == size_selected))$file
    coexpression_df = fread(coexpression_file)
    # Turn unsignificant co-expression values to 0
    #coexpression_df$CN = ifelse(coexpression_df$pval.fisher >= threshold_selected, 0, coexpression_df$CN)
    # Rename CN column into size selected
    coexpression_df = coexpression_df %>% 
      select(-pval.fisher) %>% 
      rename(!!as.character(size_selected) := "CN")
    # Merge current co-expression with other co-expression sizes
    if (i == 1){
      coexpression_merged = data.frame(coexpression_df)
    } else {
      coexpression_merged = coexpression_merged %>% left_join(coexpression_df, by=c("Node.1", "Node.2"))
    }
    rm(coexpression_df)
  }
  coexpression_merged %>% fwrite(coexpression_file_merged)
} else {
  coexpression_merged = fread(coexpression_file_merged)
}
```

```{r}
coexpression_merged %>%
  rename("20"="X20") %>%
  pivot_longer(cols=c(-Node.1, -Node.2), names_to="size", values_to="correlation", names_transform=list(size = as.character), values_transform=list(correlation = as.numeric)) %>%
  ggplot(aes(x = size, y = correlation)) +
  geom_half_violin()
```

```{r}
coexpression_file_merged = sprintf("/scratch/j.aguirreplans/Scipher/SampleSize/networks_gtex/reads/%s_%s_pearson_combined_head100000.txt", dataset, type_dataset)
networks_df = fread(coexpression_file_merged, header = TRUE)
```

```{r}
networks_df = networks_df %>% 
  mutate(gp = paste(Node.1, Node.2, sep = " -- ")) %>%
  select(!(c(Node.1, Node.2))) %>%
  group_by(gp) %>%
  pivot_longer(!gp, names_to = "ss.rep", values_to = "score") %>%
  mutate(sd=sd(score)) %>%
  ungroup() %>% 
  separate("ss.rep", into=c("ss", "rep"), sep="[.]")

networks_df$ss = as.integer(networks_df$ss)
networks_df$rep = as.integer(networks_df$rep)
max_sample_cut = plyr::round_any(max(networks_df$ss), 100, f=ceiling)
```

```{r}
networks_df %>% mutate(ssgr=cut(ss, breaks= seq(0, max_sample_cut, by = 100))) %>%
  group_by(gp, ssgr) %>%
  mutate(sdgr=sd(score)) %>%
  ungroup()
```

```{r}
coexpression_file_merged = sprintf("/scratch/j.aguirreplans/Scipher/SampleSize/networks_gtex/reads/%s_%s_pearson_combined_with_5_reps_head100000.txt", dataset, type_dataset)
networks_df = fread(coexpression_file_merged, header = TRUE)
```

Everything at once:

```{r}
networks_sd_df = networks_df %>% 
  mutate(gp = paste(Node.1, Node.2, sep = " -- ")) %>%
  select(!(c(Node.1, Node.2))) %>%
  pivot_longer(!gp, names_to = "ss.rep", values_to = "correlation") %>%
  separate("ss.rep", into=c("ss", "rep"), sep="[.]") %>%
  group_by(gp, ss) %>%
  summarise(sd_correlation=sd(correlation)) %>%
  ungroup()
```

```{r}
networks_sd_df %>%
  mutate(ss = factor(ss, levels = as.character(sort(as.numeric(unique(networks_sd_df$ss)))))) %>%
  ggplot(aes(x = ss, y = sd_correlation)) +
  geom_half_violin(side = "r", color="orangered4", fill="orangered1") +
  theme_linedraw() +
  theme(aspect.ratio=1, 
        plot.title =  element_text(size = 15, face="bold"), 
        axis.title = element_text(size = 14, face="bold"), 
        axis.text = element_text(size = 13), 
        legend.text = element_text(size = 13), 
        legend.title=element_text(size=14, face="bold")) +
  labs (x = "Number of samples",
        y = "Co-expression weight SD",
        color = "",
        fill = "")
```

Calculation in chunks:

```{r}
con <- file(coexpression_file_merged,"r")
first_line <- readLines(con,n=1)
close(con)
cols = strsplit(first_line, split=",")[[1]]

chunk_size = 50000
x=0
chunk_count = 0
connection = file(coexpression_file_merged, "r")
network_sd = data.frame()
repeat {
  x=x+1
  chunk_count=chunk_count+chunk_size
  message(paste('Chunk:', chunk_count))
  skip_param=0
  if(x==1){
    skip_param=1
  }
  network_chunk = read.table(connection, nrows=chunk_size, skip=skip_param, header=FALSE, fill = TRUE, sep=",", col.names=cols)
  network_chunk = network_chunk %>% 
    mutate(gp = paste(Node.1, Node.2, sep = " -- ")) %>%
    select(!(c(Node.1, Node.2))) %>%
    pivot_longer(!gp, names_to = "ss.rep", values_to = "correlation") %>%
    separate("ss.rep", into=c("ss", "rep"), sep="[.]") %>%
    mutate(across('ss', str_replace, 'X', '')) %>%
    group_by(gp, ss) %>%
    summarise(sd_correlation=sd(correlation)) %>%
    ungroup()
  if(nrow(network_sd) == 0){
    network_sd = data.frame(network_chunk)
  } else {
    network_sd = rbind(network_sd, network_chunk)
  }
  if(nrow(network_chunk) < chunk_size){
    close(connection)
    break
  }
  rm(network_chunk)
}
```



```{r}
network_sd %>%
  mutate(ss = factor(ss, levels = as.character(sort(as.numeric(unique(networks_sd_df$ss)))))) %>%
  ggplot(aes(x = ss, y = sd_correlation)) +
  geom_half_violin(side = "r", color="orangered4", fill="orangered1") +
  theme_linedraw() +
  theme(aspect.ratio=1, 
        plot.title =  element_text(size = 15, face="bold"), 
        axis.title = element_text(size = 14, face="bold"), 
        axis.text = element_text(size = 13), 
        legend.text = element_text(size = 13), 
        legend.title=element_text(size=14, face="bold")) +
  labs (x = "Number of samples",
        y = "Co-expression weight SD",
        color = "",
        fill = "")
```

