---
title: "Create Subsamples of gene expression data from RA patients"
author: "Joaquim Aguirre Plans"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Create subsamples of gene expression data from RA patients, with the aim of testing the effect of sample size in these subsamples.

```{r, message=FALSE}
require(data.table)
library(dplyr)
require(igraph)
require(NetSci)
```

## Read gene expression data

```{r read_gene_expression}
metadata_RNA = fread("/Users/quim/Documents/DATA/BioGroup/ScipherMedicine/data/out/Meta_RNA_Cleaned.csv")
patients <- metadata_RNA$Subject
```

The total number of patients is the dataset is `r length(patients)`.

We check the number of labels in these patients:

```{r define_response}
labels_R_3m = metadata_RNA$Subject[metadata_RNA$acr50_3m == "responder"]
labels_R_6m =metadata_RNA$Subject[metadata_RNA$acr50_3m == "nonresponder"]
labels_NR_3m = metadata_RNA$Subject[metadata_RNA$acr50_6m == "responder"]
labels_NR_6m = metadata_RNA$Subject[metadata_RNA$acr50_6m == "nonresponder"]
labels_R_3m <- paste(labels_R_3m, 'R3m', sep='_')
labels_R_6m <- paste(labels_R_6m, 'R6m', sep='_')
labels_NR_3m <- paste(labels_NR_3m, 'NR3m', sep='_')
labels_NR_6m <- paste(labels_NR_6m, 'NR6m', sep='_')
all_labels <- c(labels_R_3m, labels_R_6m, labels_NR_3m, labels_NR_6m)
```

* Responder (3M): `r length(labels_R_3m)`
* Responder (6M): `r length(labels_R_6m)`
* Non-responder (3M): `r length(labels_NR_3m)`
* Non-responder (6M): `r length(labels_NR_6m)`
* Total number of labels: `r length(length(all_labels))`

We check the balance of responder and non-responder patients in the dataset.

```{r table_response}
table(metadata_RNA %>% select(acr50_3m, acr50_6m))
```


## Subsampling

We define the sizes of the subsamples.

```{r rep_step_selection}
rep = 5
step = 10
```

We define the function to make the sampling

```{r sampling_function}
create_random_subsamples <- function(samples_list, size_list, rep){
  subsampling_cols <- c('size', 'rep', samples_list)
  subsampling_df <- data.frame(matrix(ncol=length(subsampling_cols), nrow=0))
  colnames(subsampling_df) <- subsampling_cols
  seed <- 1
  for(size in size_list){
    set.seed(seed)
    seed <- seed + 1
    for(r in 1:rep){
      #print(c(size, r))
      subsample <- sample(samples_list, size=size, replace=FALSE)
      #print(subsample)
      subsampling_df <- rbind(subsampling_df, c(size, r, ifelse(samples_list %in% subsample, 1, 0)))
      colnames(subsampling_df) <- subsampling_cols
      #sub_df <- metadata_RNA[metadata_RNA$Subject %in% subsample]
      #print(c(length(sub_df$Subject[sub_df$acr50_3m == "responder"]), length(sub_df$Subject[sub_df$acr50_3m == "nonresponder"]), length(sub_df$Subject[sub_df$acr50_6m == "responder"]), length(sub_df$Subject[sub_df$acr50_6m == "nonresponder"])))
    }
  }
  return(subsampling_df)
}
```


### Randomly sampling *n* patients

```{r size_list_selection_random_patients}
max_num_samples = length(patients)
size_list <- 1:max_num_samples
size_list <- size_list[(size_list %% step) == 0]
size_list
```

We randomly select patients for each subsample. The process is repeated `r rep` times for each subsample size.

```{r subsampling_random_patients}
random_patients_subsampling_df <- create_random_subsamples(samples_list=patients, size_list=size_list, rep=rep)
random_patients_subsampling_file <- '/Users/quim/Documents/DATA/BioGroup/ScipherMedicine/data/subsampling/subsampling_random_patients.txt'
random_patients_subsampling_df %>% write.table(random_patients_subsampling_file, quote = F, sep='\t', row.names = F, col.names = T)

# Selecting the patients for sample size 10 and repetition 1
sub_df <- random_patients_subsampling_df %>% dplyr::filter(size==10, rep==1) %>% dplyr::select(-size, -rep)
patients_sub <- colnames(sub_df[which(sub_df==1)])
```


### Randomly sampling *n* labels

```{r size_list_selection_random_labels}
max_num_samples = length(all_labels)
size_list <- 1:max_num_samples
size_list <- size_list[(size_list %% step) == 0]
size_list
```

We randomly select labels for each subsample. The process is repeated `r rep` times for each subsample size.

```{r subsampling_random_labels}
random_labels_subsampling_df <- create_random_subsamples(samples_list=all_labels, size_list=size_list, rep=rep)
random_labels_subsampling_file <- '/Users/quim/Documents/DATA/BioGroup/ScipherMedicine/data/subsampling/subsampling_random_labels.txt'
random_labels_subsampling_df %>% write.table(random_labels_subsampling_file, quote = F, sep='\t', row.names = F, col.names = T)

# Selecting the patients for sample size 10 and repetition 1
sub_df <- random_labels_subsampling_df %>% dplyr::filter(size==10, rep==1) %>% dplyr::select(-size, -rep)
labels_sub <- colnames(sub_df[which(sub_df==1)])
```

### Randomly sampling *n* labels maintaining the original proportion

We define the function to make the sampling

```{r subsampling_random_labels_of_specific_proportion}
## Stratified bootstrapping
create_random_subsamples_of_specific_proportion <- function(size_list, rep, label_to_patients_list, label_to_proportion_list){
  samples_list <- as.vector(unlist(label_to_patients_list))
  subsampling_cols <- c('size', 'rep', samples_list)
  subsampling_df <- data.frame(matrix(ncol=length(subsampling_cols), nrow=0))
  colnames(subsampling_df) <- subsampling_cols
  seed <- 1
  for(size in size_list){
    set.seed(seed)
    seed <- seed + 1
    for(r in 1:rep){
      #print(c(size, r))
      samples_selection <- c()
      for(label in names(label_to_patients_list)){
        patients <- label_to_patients_list[[label]]
        proportion <- label_to_proportion_list[[label]]
        subsample <- sample(patients, size=size*proportion, replace=FALSE)
        #print(subsample)
        samples_selection <- c(samples_selection, subsample)
      }
      subsampling_df <- rbind(subsampling_df, c(size, r, ifelse(samples_list %in% samples_selection, 1, 0)))
      colnames(subsampling_df) <- subsampling_cols
      #sub_df <- metadata_RNA[metadata_RNA$Subject %in% subsample]
      #print(c(length(sub_df$Subject[sub_df$acr50_3m == "responder"]), length(sub_df$Subject[sub_df$acr50_3m == "nonresponder"]), length(sub_df$Subject[sub_df$acr50_6m == "responder"]), length(sub_df$Subject[sub_df$acr50_6m == "nonresponder"])))
    }
  }
  return(subsampling_df)
}

select_size_list_by_label_proportion <- function(max_num_samples, step, label_to_patients_list, label_to_proportion_list){
  initial_size_list <- 1:max_num_samples
  initial_size_list <- initial_size_list[(initial_size_list %% step) == 0]
  size_list <- c()
  for(size in initial_size_list){
    size_ok = TRUE
    for(label in names(label_to_patients_list)){
      patients <- label_to_patients_list[[label]]
      proportion <- label_to_proportion_list[[label]]
      if(length(patients) < size*proportion){
        size_ok = FALSE
      }
    }
    if(!size_ok){
      break
    } else {
      size_list <- c(size_list, size)
    }
  }
  return(size_list)
}

```

We randomly select labels for each subsample. The process is repeated `r rep` times for each subsample size.

```{r size_list_selection_random_labels_of_original_proportion}
max_num_samples = length(all_labels)
label_to_patients_list <- list(R3m = labels_R_3m, R6m = labels_R_6m, NR3m = labels_NR_3m, NR6m = labels_NR_6m)
label_to_proportion_list <- list(R3m = length(labels_R_3m)/length(all_labels), R6m = length(labels_R_6m)/length(all_labels), NR3m = length(labels_NR_3m)/length(all_labels), NR6m = length(labels_NR_6m)/length(all_labels))
initial_size_list <- 1:max_num_samples
initial_size_list <- initial_size_list[(initial_size_list %% step) == 0]
initial_size_list
size_list <- select_size_list_by_label_proportion(max_num_samples, step, label_to_patients_list, label_to_proportion_list)
size_list

```

```{r subsampling_random_labels_of_original_proportion}
random_labels_subsampling_original_proportion_df <- create_random_subsamples_of_specific_proportion(size_list=size_list, rep=rep, label_to_patients_list=label_to_patients_list, label_to_proportion_list=label_to_proportion_list)
random_labels_subsampling_original_proportion_file <- '/Users/quim/Documents/DATA/BioGroup/ScipherMedicine/data/subsampling/subsampling_random_labels_original_proportion.txt'
random_labels_subsampling_original_proportion_df %>% write.table(random_labels_subsampling_original_proportion_file, quote = F, sep='\t', row.names = F, col.names = T)

```


### Randomly sampling *n* labels maintaining an equal proportion

```{r size_list_selection_random_labels_of_equal_proportion}
max_num_samples = length(all_labels)
label_to_patients_list <- list(R3m = labels_R_3m, R6m = labels_R_6m, NR3m = labels_NR_3m, NR6m = labels_NR_6m)
label_to_proportion_list <- list(R3m = 0.25, R6m = 0.25, NR3m = 0.25, NR6m = 0.25)
initial_size_list <- 1:max_num_samples
initial_size_list <- initial_size_list[(initial_size_list %% step) == 0]
initial_size_list
size_list <- select_size_list_by_label_proportion(max_num_samples, step, label_to_patients_list, label_to_proportion_list)
size_list

```

```{r subsampling_random_labels_of_original_proportion}
random_labels_subsampling_equal_proportion_df <- create_random_subsamples_of_specific_proportion(size_list=size_list, rep=rep, label_to_patients_list=label_to_patients_list, label_to_proportion_list=label_to_proportion_list)
random_labels_subsampling_equal_proportion_file <- '/Users/quim/Documents/DATA/BioGroup/ScipherMedicine/data/subsampling/subsampling_random_labels_equal_proportion.txt'
random_labels_subsampling_equal_proportion_df %>% write.table(random_labels_subsampling_equal_proportion_file, quote = F, sep='\t', row.names = F, col.names = T)

# Selecting the patients for sample size 10 and repetition 1
sub_df <- random_labels_subsampling_equal_proportion_df %>% dplyr::filter(size==10, rep==1) %>% dplyr::select(-size, -rep)
patients_sub <- colnames(sub_df[which(sub_df==1)])

```
