---
title: "Analyze differentially co-expressed genes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Analyze differentially co-expressed genes.

Based on this tutorial:
[https://deisygysi.github.io/teaching/minipeb2021/](https://deisygysi.github.io/teaching/minipeb2021/)

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(igraph)
library(ggalluvial)
library(ggplot2)
require(magrittr)
library(tidyr)
set.seed(1510)
options(bitmapType='cairo')
`%ni%` <- Negate(`%in%`)
```

### Load data

```{r}
plots_dir = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots"
tables_dir = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/tables"
```

### Prepare disease genes information dataset

Prepare the disease-gene associations: 

```{r}
disease_genes_info_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/disease_genes/disease_genes_info_2022.csv"

if(!(file.exists(disease_genes_info_file))){
  gene_associations_dir = '/work/ccnr/j.aguirreplans/data/gene_associations/data/out'
  disease_genes_file = paste(gene_associations_dir, 'GDA_Filtered_04042022.csv', sep='/')
  GDA = fread(disease_genes_file) %>% unique()
  GDA$HGNC_Symbol = toupper(GDA$HGNC_Symbol)
  GDA[is.na(GDA)]<-0
  GDA$NewName = tolower(GDA$NewName)
  GDA$DescriptorName = tolower(GDA$DescriptorName)
  GDA$MESHID = toupper(GDA$MESHID)
  GDA$HGNC_Symbol = toupper(GDA$HGNC_Symbol)
  GDA %<>% 
    filter(Strong > 0 | 
             Weak > 0 |
             Incompatible > 0) %>%
    mutate(DiseaseName = NewName) %>%
    filter(DescriptorName %ni% c("behavioral disciplines and activities")) %>%
    mutate(MESHID_remove = stringr::str_detect(TreeNumber, pattern = "C22|C25|F04")) %>%
    mutate(MESHID_remove = ifelse(TreeNumber == "F01", TRUE,MESHID_remove)) %>%
    filter(!(MESHID_remove)) %>%
    select(DiseaseName, HGNC_Symbol, DescriptorName) %>% 
    unique()
  
  # Make columns without special characters  
  GDA$DiseaseName.no.sp.char <- gsub(' ', '.', gsub(', ', '.', gsub('-', '.', gsub('[\\(\\)]', '', GDA$DiseaseName))))
  GDA$DescriptorName.no.sp.char <- gsub(' ', '.', gsub(', ', '.', gsub('-', '.', gsub('[\\(\\)]', '', GDA$DescriptorName))))
  
  # Write output file
  GDA %>% fwrite(disease_genes_info_file)
} else {
  GDA = fread(disease_genes_info_file)
}

head(GDA)
```

There are `r nrow(GDA)` disease-gene associations (`r length(unique(GDA$DiseaseName))` diseases, `r length(unique(GDA$HGNC_Symbol))` genes).


### Analyze breast carcinoma

Filter disease-gene associations by the disease "breast neoplasms":

```{r}
disease_name = "breast.neoplasms"
GDA_breast = GDA %>% select("DiseaseName.no.sp.char", "HGNC_Symbol") %>% filter(DiseaseName.no.sp.char == disease_name)
dim(GDA_breast)
head(GDA_breast)
```

Compile the files of the differential co-expression analysis from CODINA classifying the nodes:

```{r}
input_dir = "/scratch/j.aguirreplans/Scipher/SampleSize/differential_coexpression_analysis/reads/TCGA-BRCA_female___Breast.Mammary.Tissue_female"
result_files = list.files(input_dir)

cols = c("file_name", "size", "rep_D", "rep_N")
file_info_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))

for(file_name in result_files){
  info_file = gsub(".txt","",file_name)
  file_split1 = strsplit(info_file, split="_")[[1]]
  if(file_split1[1] == "diffanalysis"){
    type_analysis = file_split1[2]
    if(type_analysis == "nodes"){
      pval = tail(file_split1, n=1)
      info_file = gsub(paste("diffanalysis_", type_analysis, "_", sep=""),"",info_file)
      info_file = gsub(paste("_pval_", pval, sep=""),"",info_file)
      file_split2 = strsplit(info_file, split="___")[[1]]
      file_D_split = strsplit(gsub(".net","",file_split2[1]), split="_")[[1]]
      file_N_split = strsplit(gsub(".net","",file_split2[2]), split="_")[[1]]
      method = file_D_split[1]
      size = as.numeric(file_D_split[(length(file_D_split)-2)])
      r_D = as.numeric(file_D_split[(length(file_D_split))])
      r_N = as.numeric(file_N_split[(length(file_N_split))])
      dataset_D = file_D_split[(length(file_D_split)-4)]
      dataset_N = file_N_split[(length(file_N_split)-4)]
      file_info_df <- rbind(file_info_df, data.frame(file_name=file_name, size=size, rep_D=r_D, rep_N=r_N))
    }
  }
}

head(file_info_df)
```

Read the files of the differential co-expression analysis from CODINA classifying the nodes, and put them together in a single table:

```{r}
pval_correction_field = "pval_Phi_Tilde.adj.fdr"
pval_threshold = 0.05
cols = c("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval", "size", "rep_D", "rep_N")
nodes_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))
missing_files = c()
sizes_list = sort(unique(file_info_df$size))

for(x in seq(sizes_list)){
  size = sizes_list[x]
  #print(size)
  same_size_files_df = file_info_df %>% filter(size==!!size) %>% arrange(rep_D, rep_N)
  input_files = same_size_files_df$file_name
  # same_size_info_df = data.frame()
  for(input_file in input_files){
    r_D = (same_size_files_df %>% filter(file_name == input_file))$rep_D
    r_N = (same_size_files_df %>% filter(file_name == input_file))$rep_N
    individual_df = fread(paste(input_dir, input_file, sep="/"))
    if (nrow(individual_df) > 0){
      node_results_filt_df = individual_df %>%
        select(Node, Phi_tilde, !!as.symbol(pval_correction_field)) %>% 
        unique() %>% 
        rename("pval" = !!as.symbol(pval_correction_field)) %>% 
        left_join(GDA_breast, by=c("Node"="HGNC_Symbol"))
      if(nrow(nodes_df) == 0){
        nodes_df = cbind((node_results_filt_df %>% select("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval")), data.frame(size=size, rep_D=r_D, rep_N=r_N))
      } else {
        nodes_df = nodes_df %>% full_join(cbind((node_results_filt_df %>% select("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval")), data.frame(size=size, rep_D=r_D, rep_N=r_N)), by=c("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval", "size", "rep_D", "rep_N"))
      }
    } else {
      missing_files = c(missing_files, input_file)
    }
  }
}

# Define Phi_name
nodes_df$Phi_name = ifelse(nodes_df$Phi_tilde == "a", "common", ifelse(nodes_df$Phi_tilde == "b.D", "different", ifelse(nodes_df$Phi_tilde == "g.D", "disease-specific", ifelse(nodes_df$Phi_tilde == "g.N", "normal-specific", "undefined"))))

head(nodes_df)
```

We consider as "undefined" both the genes that are classified as U, but also the ones that are not significantly classified in any of the categories. To do so, we need create an empty table with all considered genes for all sizes and repetitions, and fill it with the data that we already have:

```{r}
reps = sort(unique(paste(nodes_df$size, nodes_df$rep_D, nodes_df$rep_N, sep="-")))
nodes_empty_df = expand.grid(sort(unique(nodes_df$Node)), reps)
nodes_empty_df = nodes_empty_df %>% separate("Var2", into=c("size", "rep_D", "rep_N"), sep="-", convert=TRUE)
colnames(nodes_empty_df) = c("Node", "size", "rep_D", "rep_N")
nodes_empty_df = nodes_empty_df %>% left_join((nodes_df %>% select(Node, DiseaseName.no.sp.char) %>% unique()), by="Node") %>% arrange(Node, size, rep_D, rep_N)
nodes_expanded_df = nodes_df %>% full_join(nodes_empty_df, by = c("Node", "DiseaseName.no.sp.char", "size", "rep_D", "rep_N")) %>% arrange(Node, size, rep_D)
nodes_expanded_df = nodes_expanded_df %>%
  mutate(Phi_tilde = replace(Phi_tilde, is.na(Phi_tilde), "U")) %>%
  mutate(Phi_name = replace(Phi_name, is.na(Phi_name), "undefined"))
nodes_expanded_file = paste(tables_dir, "codina_tcga_brca_gtex_breast_female_nodes.txt", sep="/")
nodes_expanded_df %>% fwrite(nodes_expanded_file)
head(nodes_expanded_df)
```

We count how many times each gene category appears in each sample size:

```{r}
# Count for each gene and size the number of times that each category is predicted
nodes_counted_df = nodes_expanded_df %>% 
  group_by(Node, DiseaseName.no.sp.char, Phi_tilde, Phi_name, size) %>% 
  summarize(n_phi = n()) %>% 
  ungroup()  %>% 
  arrange(Node, size)

nodes_counted_file = paste(tables_dir, "codina_tcga_brca_gtex_breast_female_nodes_counted.txt", sep="/")
nodes_counted_df %>% fwrite(nodes_counted_file)
head(nodes_counted_df)
```

We select the gene category that appears more in the different repetitions of each sample size:

```{r}
# Select the for each gene and size the node type that occurs more times in the different repetitions
nodes_filtered_df = nodes_counted_df %>%
  group_by(Node, DiseaseName.no.sp.char, size) %>%
  filter(n_phi == max(n_phi)) %>% 
  arrange(Node, size) %>% 
  sample_n(1) %>% # Choose randomly cases in which there is the same number of node types 
  ungroup()
 
nodes_filtered_file = paste(tables_dir, "codina_tcga_brca_gtex_breast_female_nodes_filtered.txt", sep="/")
nodes_filtered_df %>% fwrite(nodes_filtered_file)
head(nodes_filtered_df)
```

We count the number of disease genes in each category for each size:

```{r}
counts_categories_empty_df = expand.grid(sort(unique(nodes_expanded_df$Phi_name)), reps)
counts_categories_empty_df = counts_categories_empty_df %>% separate("Var2", into=c("size", "rep_D", "rep_N"), sep="-", convert=TRUE)
colnames(counts_categories_empty_df) = c("Phi_name", "size", "rep_D", "rep_N")
counts_categories_df = nodes_expanded_df %>% 
  filter(!(is.na(DiseaseName.no.sp.char))) %>% 
  group_by(Phi_name, size, rep_D, rep_N) %>% count() %>% rename("n_disease"="n") %>% arrange(size)
counts_categories_df = counts_categories_df %>% full_join(counts_categories_empty_df, by = c("Phi_name", "size", "rep_D", "rep_N")) %>% arrange(Phi_name, size, rep_D, rep_N)
counts_categories_df$n_disease = ifelse(is.na(counts_categories_df$n_disease), 0, counts_categories_df$n_disease)
counts_categories_file = paste(tables_dir, "codina_tcga_brca_gtex_breast_female_category_counts.txt", sep="/")
counts_categories_df %>% fwrite(counts_categories_file)
head(counts_categories_df)
```

We plot the number of disease genes in each category:

```{r}
custom_palette = c(
                   "#619CFF", #common
                   "#F8766D", #disease-specific 
                   "#00BA38", #normal-specific
                   "#808080", #undefined
                   "#C77CFF" #different
                   )

counts_categories_df %>% 
  group_by(Phi_name, size) %>%
  summarize(mean_disease = mean(n_disease), sd_disease = sd(n_disease), max_disease=max(n_disease), min_disease=min(n_disease)) %>%
  ggplot(aes(x = size, y = mean_disease, fill = Phi_name)) + 
  geom_line(aes(col = Phi_name)) +
  geom_ribbon(aes(ymin=min_disease, ymax=max_disease), alpha=0.2) +
  xlab("Number of samples") +
  ylab("Number of disease genes") +
  scale_fill_manual(values = custom_palette) +
  scale_color_manual(values = custom_palette) +
  guides(col=guide_legend(title="Gene category"), fill=guide_legend(title="Gene category")) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 14, face="bold"), axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title=element_text(size=14, face="bold"))
plot_file = paste(plots_dir, "codina_tcga_brca_gtex_breast_female_counts_disease_genes.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```

We create the alluvium plot showing the flow between gene categories across sample size:

```{r}
nodes_filtered_df$Phi_name <- as.factor(nodes_filtered_df$Phi_name)
nodes_filtered_df$size = as.character(nodes_filtered_df$size)
levels(nodes_filtered_df$Phi_name) = sort(unique(nodes_filtered_df$Phi_name))
nodes_filtered_df$size <- factor(nodes_filtered_df$size, levels=as.character(sort(unique(as.integer(nodes_filtered_df$size)))))

custom_palette = c(
                   "#619CFF", #common
                   "#F8766D", #disease-specific 
                   "#00BA38", #normal-specific
                   "#808080", #undefined
                   "#C77CFF" #different
                   )
plot_file = paste(plots_dir, "codina_tcga_brca_gtex_breast_female_change_disease_genes.png", sep="/")
nodes_filtered_df %>% 
  filter(size %in% c(20, 80, 140, 200, 260, 320, 380, 440)) %>%
  filter(!(is.na(DiseaseName.no.sp.char))) %>%
  ggplot(aes(x = size, stratum = Phi_name, alluvium = Node, 
             fill = Phi_name, label = Phi_name)) +
  scale_fill_manual(values = custom_palette) +
  scale_color_manual(values = custom_palette) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback") +
  geom_stratum() +
  xlab("Number of samples") +
  ylab("Number of disease genes") +
  guides(fill=guide_legend(title="Gene category")) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), 
        axis.title = element_text(size = 14, face="bold"), 
        axis.text = element_text(size = 12), 
        #legend.position = "bottom",
        legend.text = element_text(size = 12), 
        legend.title=element_text(size=14, face="bold"))
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```

We check the behavior specific genes associated with breast neoplasms:

```{r}
# Nodes of breast neoplasms: https://www.nationalbreastcancer.org/other-breast-cancer-genes
# https://www.cancer.org/cancer/breast-cancer/risk-and-prevention/breast-cancer-risk-factors-you-cannot-change.html#:~:text=BRCA1%20and%20BRCA2%3A%20The%20most,which%20can%20lead%20to%20cancer.
nodes_to_follow = c("BRCA1", "BRCA2", "PALB2", "CHEK2", "CDH1", "PTEN", "STK11", "TP53") ### WE ARE MISSING BRCA2!!!!
for (important_node in nodes_to_follow){
  # Filter results by important node
  important_df = nodes_expanded_df %>% 
    filter(size %in% c(20, 80, 140, 200, 260, 320, 380, 440)) %>%
    filter(!(is.na(DiseaseName.no.sp.char))) %>%
    filter(Node == important_node) %>%
    select(Phi_name, size, rep_D, rep_N) %>%
    group_by(Phi_name, size) %>% 
    summarize(n_evidences = n()) %>% 
    ungroup()
  # Define levels
  important_df$Phi_name <- as.factor(important_df$Phi_name)
  important_df$size = as.character(important_df$size)
  levels(important_df$Phi_name) = sort(unique(important_df$Phi_name))
  important_df$size <- factor(important_df$size, levels=as.character(sort(unique(as.integer(important_df$size)))))
  # Define palette
  if (levels(important_df$Phi_name) == c("disease-specific", "normal-specific", "undefined")){
    personalized_palette = c("#F8766D", "#00BA38", "#808080")
  } else if (levels(important_df$Phi_name) == c("normal-specific", "undefined")){
    personalized_palette = c("#00BA38", "#808080")
  } else if (levels(important_df$Phi_name) == c("common", "normal-specific", "undefined")){
    personalized_palette = c("#619CFF", "#00BA38", "#808080")
  } else if (levels(important_df$Phi_name) == c("normal-specific")){
    personalized_palette = c("#00BA38")
  } else {
    personalized_palette = c("#619CFF", "#F8766D", "#00BA38", "#808080", "#C77CFF")
  }
  # Plot
  important_node_plot = important_df %>%
    ggplot(aes(fill=Phi_name, y=n_evidences, x=size)) + 
      geom_bar(position="fill", stat="identity", width = 0.7, col="black") +
      #geom_bar(position = position_dodge(), stat="identity", col="black", width = 0.6) +
    xlab("Number of samples") +
    ylab("Fraction of evidences") +
    scale_fill_manual(values = personalized_palette) +
    guides(fill=guide_legend(title="Gene category")) +
    theme_linedraw() +
    theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 14, face="bold"), axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title=element_text(size=14, face="bold"))
  plot_file = paste(plots_dir, paste("codina_tcga_brca_gtex_breast_female_important_", important_node, ".png", sep=""), sep="/")
  ggsave(
    plot_file,
    plot=important_node_plot,
    dpi = 1200,
    width = 9000,
    height = 6000,
    units = c("px")
  )
  print(important_node_plot)
}
```

Calculate the fraction of unchanged / changed genes in each category for each sample size:

```{r}
# Define output table
cols = c("Node", "Phi_name.prev", "size.prev", "Phi_name.curr", "size.curr")
change_of_category_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))
# Remove levels
nodes_filtered_df$size = as.numeric(as.character(nodes_filtered_df$size))
# For each size, compare nodes with the ones of previous size
for (x in seq(sizes_list)){
  size = sizes_list[x]
  nodes_filtered_by_selected_size = nodes_filtered_df %>%
    filter((!(is.na(DiseaseName.no.sp.char))) & (size == !!size)) %>%
    select(Node, Phi_name, size)
  if (x == 1){
    previous_nodes_filtered = data.frame(nodes_filtered_by_selected_size)
  } else {
    nodes_filtered_compared_by_sizes = previous_nodes_filtered %>%
      inner_join(nodes_filtered_by_selected_size, by=("Node")) %>%
      rename(c("Phi_name.prev"="Phi_name.x", "Phi_name.curr"="Phi_name.y", "size.prev"="size.x", "size.curr"="size.y"))
    change_of_category_df = rbind(change_of_category_df, nodes_filtered_compared_by_sizes)
    previous_nodes_filtered = data.frame(nodes_filtered_by_selected_size)
  }
}

# Detect changes
change_of_category_df$transition_type = ifelse(change_of_category_df$Phi_name.prev == change_of_category_df$Phi_name.curr, "stable", "change")
change_of_category_df$transition_name = paste(change_of_category_df$Phi_name.prev, change_of_category_df$Phi_name.curr, sep=" / ")

# Print table
change_of_category_file = paste(tables_dir, "codina_tcga_brca_gtex_breast_changes.txt", sep="/")
change_of_category_df %>% fwrite(change_of_category_file)
head(change_of_category_df)
```

```{r}
total_stable_change_genes_df = change_of_category_df %>%
  group_by(size.prev) %>% 
  mutate(n_total = n()) %>%
  ungroup() %>%
  select(Node, Phi_name.prev, size.prev, transition_type, n_total) %>% 
  group_by(size.prev, transition_type, n_total) %>% 
  summarize(n_genes = n()) %>% 
  ungroup() %>%
  mutate(frac_genes = n_genes / n_total)
total_stable_change_genes_df$Phi_name.prev = "all"

categories_stable_change_genes_df = change_of_category_df %>%
  group_by(size.prev) %>% 
  mutate(n_total = n()) %>%
  ungroup() %>%
  filter(transition_type == "stable") %>%
  select(Node, Phi_name.prev, size.prev, transition_type, n_total) %>% 
  group_by(Phi_name.prev, size.prev, transition_type, n_total) %>% 
  summarize(n_genes = n()) %>% 
  ungroup() %>%
  mutate(frac_genes = n_genes / n_total)

rbind(total_stable_change_genes_df, categories_stable_change_genes_df) %>%
  filter(transition_type == "stable") %>%
  ggplot(aes(x = size.prev, y = frac_genes, col = Phi_name.prev)) + 
  geom_line(aes(size = Phi_name.prev)) +
  xlab("Number of samples") +
  ylab("Fraction of stable genes") +
  scale_size_manual(values = c("all" = 1, "common" = 0.5, "disease-specific" = 0.5, "normal-specific" = 0.5, "undefined" = 0.5)) +
  scale_color_manual(values = c("#CD9600", "#619CFF", "#F8766D", "#00BA38", "#808080")) +
  guides(col=guide_legend(title="Gene category"), size=FALSE) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 14, face="bold"), axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title=element_text(size=14, face="bold"))
plot_file = paste(plots_dir, "codina_tcga_brca_gtex_breast_stable_genes.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```



### Analyze rheumatoid arthritis

Filter disease-gene associations by the disease "arthritis rheumatoid":

```{r}
disease_name = "arthritis.rheumatoid"
GDA_ra = GDA %>% select("DiseaseName.no.sp.char", "HGNC_Symbol") %>% filter(DiseaseName.no.sp.char == disease_name)
dim(GDA_ra)
head(GDA_ra)
```

Compile the files of the differential co-expression analysis from CODINA:

```{r}
input_dir = "/work/ccnr/j.aguirreplans/Scipher/SampleSize/differential_coexpression_analysis/complete.dataset_Whole.Blood"
result_files = list.files(input_dir)

cols = c("file_name", "size", "rep_D", "rep_N")
file_info_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))

for(file_name in result_files){
  info_file = gsub(".txt","",file_name)
  file_split1 = strsplit(info_file, split="_")[[1]]
  if(file_split1[1] == "diffanalysis"){
    type_analysis = file_split1[2]
    if(type_analysis == "nodes"){
      pval = tail(file_split1, n=1)
      info_file = gsub(paste("diffanalysis_", type_analysis, "_", sep=""),"",info_file)
      info_file = gsub(paste("_pval_", pval, sep=""),"",info_file)
      file_split2 = strsplit(info_file, split="___")[[1]]
      file_D_split = strsplit(gsub(".net","",file_split2[1]), split="_")[[1]]
      file_N_split = strsplit(gsub(".net","",file_split2[2]), split="_")[[1]]
      method = file_D_split[1]
      size = as.numeric(file_D_split[(length(file_D_split)-2)])
      r_D = as.numeric(file_D_split[(length(file_D_split))])
      r_N = as.numeric(file_N_split[(length(file_N_split))])
      dataset_D = file_D_split[(length(file_D_split)-4)]
      dataset_N = file_N_split[(length(file_N_split)-4)]
      file_info_df <- rbind(file_info_df, data.frame(file_name=file_name, size=size, rep_D=r_D, rep_N=r_N))
    }
  }
}

head(file_info_df)
```

Read the files of the differential co-expression analysis from CODINA classifying the nodes, and put them together in a single table:

```{r}
pval_correction_field = "pval_Phi_Tilde.adj.fdr"
pval_threshold = 0.05
cols = c("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval", "size", "rep_D", "rep_N")
nodes_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))
sizes_list = sort(unique(file_info_df$size))

for(x in seq(sizes_list)){
  size = sizes_list[x]
  #print(size)
  same_size_files_df = file_info_df %>% filter(size==!!size) %>% arrange(rep_D, rep_N)
  input_files = same_size_files_df$file_name
  # same_size_info_df = data.frame()
  for(input_file in input_files){
    r_D = (same_size_files_df %>% filter(file_name == input_file))$rep_D
    r_N = (same_size_files_df %>% filter(file_name == input_file))$rep_N
    individual_df = fread(paste(input_dir, input_file, sep="/"))
    node_results_filt_df = individual_df %>%
      select(Node, Phi_tilde, !!as.symbol(pval_correction_field)) %>% 
      unique() %>% 
      rename("pval" = !!as.symbol(pval_correction_field)) %>% 
      left_join(GDA_ra, by=c("Node"="HGNC_Symbol"))
    if(nrow(nodes_df) == 0){
      nodes_df = cbind((node_results_filt_df %>% select("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval")), data.frame(size=size, rep_D=r_D, rep_N=r_N))
    } else {
      nodes_df = nodes_df %>% full_join(cbind((node_results_filt_df %>% select("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval")), data.frame(size=size, rep_D=r_D, rep_N=r_N)), by=c("Node", "DiseaseName.no.sp.char", "Phi_tilde", "pval", "size", "rep_D", "rep_N"))
    }
  }
}

# Define Phi_name
nodes_df$Phi_name = ifelse(nodes_df$Phi_tilde == "a", "common", ifelse(nodes_df$Phi_tilde == "b.D", "different", ifelse(nodes_df$Phi_tilde == "g.D", "disease-specific", ifelse(nodes_df$Phi_tilde == "g.N", "normal-specific", "undefined"))))

head(nodes_df)
```

We consider as "undefined" both the genes that are classified as U, but also the ones that are not significantly classified in any of the categories. To do so, we need create an empty table with all considered genes for all sizes and repetitions, and fill it with the data that we already have:

```{r}
reps = sort(unique(paste(nodes_df$size, nodes_df$rep_D, nodes_df$rep_N, sep="-")))
nodes_empty_df = expand.grid(sort(unique(nodes_df$Node)), reps)
nodes_empty_df = nodes_empty_df %>% separate("Var2", into=c("size", "rep_D", "rep_N"), sep="-", convert=TRUE)
colnames(nodes_empty_df) = c("Node", "size", "rep_D", "rep_N")
nodes_empty_df = nodes_empty_df %>% left_join((nodes_df %>% select(Node, DiseaseName.no.sp.char) %>% unique()), by="Node") %>% arrange(Node, size, rep_D, rep_N)
nodes_expanded_df = nodes_df %>% full_join(nodes_empty_df, by = c("Node", "DiseaseName.no.sp.char", "size", "rep_D", "rep_N")) %>% arrange(Node, size, rep_D)
nodes_expanded_df = nodes_expanded_df %>%
  mutate(Phi_tilde = replace(Phi_tilde, is.na(Phi_tilde), "U")) %>%
  mutate(Phi_name = replace(Phi_name, is.na(Phi_name), "undefined"))
nodes_expanded_file = paste(tables_dir, "codina_scipher_ra_gtex_wholeblood_nodes.txt", sep="/")
nodes_expanded_df %>% fwrite(nodes_expanded_file)
head(nodes_expanded_df)
```

We count how many times each gene category appears in each sample size:

```{r}
# Count for each gene and size the number of times that each category is predicted
nodes_counted_df = nodes_expanded_df %>% 
  group_by(Node, DiseaseName.no.sp.char, Phi_tilde, Phi_name, size) %>% 
  summarize(n_phi = n()) %>% 
  ungroup()  %>% 
  arrange(Node, size)

nodes_counted_file = paste(tables_dir, "codina_scipher_ra_gtex_wholeblood_nodes_counted.txt", sep="/")
nodes_counted_df %>% fwrite(nodes_counted_file)
head(nodes_counted_df)
```

We select the gene category that appears more for in the different repetitions of each sample size:

```{r}
# Select the for each gene and size the node type that occurs more times in the different repetitions
nodes_filtered_df = nodes_counted_df %>%
  group_by(Node, DiseaseName.no.sp.char, size) %>%
  filter(n_phi == max(n_phi)) %>% 
  arrange(Node, size) %>% 
  sample_n(1) %>% # Choose randomly cases in which there is the same number of node types 
  ungroup()
 
nodes_filtered_file = paste(tables_dir, "codina_scipher_ra_gtex_wholeblood_nodes_filtered.txt", sep="/")
nodes_filtered_df %>% fwrite(nodes_filtered_file)
head(nodes_filtered_df)
```

We count the number of disease genes in each category for each size:

```{r}
counts_categories_empty_df = expand.grid(sort(unique(nodes_expanded_df$Phi_name)), reps)
counts_categories_empty_df = counts_categories_empty_df %>% separate("Var2", into=c("size", "rep_D", "rep_N"), sep="-", convert=TRUE)
colnames(counts_categories_empty_df) = c("Phi_name", "size", "rep_D", "rep_N")
counts_categories_df = nodes_expanded_df %>% 
  filter(!(is.na(DiseaseName.no.sp.char))) %>% 
  group_by(Phi_name, size, rep_D, rep_N) %>% count() %>% rename("n_disease"="n") %>% arrange(size)
counts_categories_df = counts_categories_df %>% full_join(counts_categories_empty_df, by = c("Phi_name", "size", "rep_D", "rep_N")) %>% arrange(Phi_name, size, rep_D, rep_N)
counts_categories_df$n_disease = ifelse(is.na(counts_categories_df$n_disease), 0, counts_categories_df$n_disease)
counts_categories_file = paste(tables_dir, "codina_scipher_ra_gtex_wholeblood_category_counts.txt", sep="/")
counts_categories_df %>% fwrite(counts_categories_file)
head(counts_categories_df)
```

We plot the number of disease genes in each category:

```{r}
custom_palette = c(
                   "#619CFF", #common
                   "#C77CFF", #different
                   "#F8766D", #disease-specific 
                   "#00BA38", #normal-specific
                   "#808080" #undefined
                   )

counts_categories_df %>% 
  group_by(Phi_name, size) %>%
  summarize(mean_disease = mean(n_disease), sd_disease = sd(n_disease), max_disease=max(n_disease), min_disease=min(n_disease)) %>%
  ggplot(aes(x = size, y = mean_disease, fill = Phi_name)) + 
  geom_line(aes(col = Phi_name)) +
  geom_ribbon(aes(ymin=min_disease, ymax=max_disease), alpha=0.2) +
  xlab("Number of samples") +
  ylab("Number of disease genes") +
  scale_fill_manual(values = custom_palette) +
  scale_color_manual(values = custom_palette) +
  guides(col=guide_legend(title="Gene category"), fill=guide_legend(title="Gene category")) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 14, face="bold"), axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title=element_text(size=14, face="bold"))
plot_file = paste(plots_dir, "codina_scipher_ra_gtex_wholeblood_counts_disease_genes.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```

We create the alluvium plot showing the flow between gene categories across sample size:

```{r}
nodes_filtered_df$Phi_name <- as.factor(nodes_filtered_df$Phi_name)
nodes_filtered_df$size = as.character(nodes_filtered_df$size)
levels(nodes_filtered_df$Phi_name) = sort(unique(nodes_filtered_df$Phi_name))
nodes_filtered_df$size <- factor(nodes_filtered_df$size, levels=as.character(sort(unique(as.integer(nodes_filtered_df$size)))))

custom_palette = c(
                   "#619CFF", #common
                   "#F8766D", #disease-specific 
                   "#00BA38", #normal-specific
                   "#808080", #undefined
                   "#C77CFF" #different
                   )
plot_file = paste(plots_dir, "codina_scipher_ra_gtex_wholeblood_change_disease_genes.png", sep="/")
nodes_filtered_df %>% 
  filter(size %in% c(20, 80, 140, 200, 260, 320, 380, 440)) %>%
  filter(!(is.na(DiseaseName.no.sp.char))) %>%
  ggplot(aes(x = size, stratum = Phi_name, alluvium = Node, 
             fill = Phi_name, label = Phi_name)) +
  scale_fill_manual(values = custom_palette) +
  scale_color_manual(values = custom_palette) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback") +
  geom_stratum() +
  xlab("Number of samples") +
  ylab("Number of disease genes") +
  guides(fill=guide_legend(title="Gene category")) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), 
        axis.title = element_text(size = 14, face="bold"), 
        axis.text = element_text(size = 12), 
        #legend.position = "bottom",
        legend.text = element_text(size = 12), 
        legend.title=element_text(size=14, face="bold"))
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```

We check the behavior specific genes associated with rheumatoid arthritis:

```{r}
# Get only strong evidences: "TNF", "IL6ST", "IL6R", "PTPN22", "HOTAIR", "MALAT1"
nodes_to_follow = c("IL6ST", "IL6R", "PTPN22", "HOTAIR", "MALAT1") #### !!!! MISSING TNF !!!!
#nodes_to_follow = c("HLA-DRB1", "HLA-DPB1", "IRF5", "PTPN22", "RBPJ", "RUNX1", "STAT4", "TNF")
for (important_node in nodes_to_follow){
  # Filter results by important node
  important_df = nodes_expanded_df %>% 
    filter(size %in% c(20, 80, 140, 200, 260, 320, 380, 440)) %>%
    filter(!(is.na(DiseaseName.no.sp.char))) %>%
    filter(Node == important_node) %>%
    select(Phi_name, size, rep_D, rep_N) %>%
    group_by(Phi_name, size) %>% 
    summarize(n_evidences = n()) %>% 
    ungroup()
  # Define levels
  important_df$Phi_name <- as.factor(important_df$Phi_name)
  important_df$size = as.character(important_df$size)
  levels(important_df$Phi_name) = sort(unique(important_df$Phi_name))
  important_df$size <- factor(important_df$size, levels=as.character(sort(unique(as.integer(important_df$size)))))
  # Define palette
  if ((length(levels(important_df$Phi_name)) == length(c("disease-specific", "normal-specific", "undefined"))) && (all(levels(important_df$Phi_name) == c("disease-specific", "normal-specific", "undefined")))){
    personalized_palette = c("#F8766D", "#00BA38", "#808080")
  } else if ((length(levels(important_df$Phi_name)) == length(c("normal-specific", "undefined"))) && (all(levels(important_df$Phi_name) == c("normal-specific", "undefined")))){
    personalized_palette = c("#00BA38", "#808080")
  } else if ((length(levels(important_df$Phi_name)) == length(c("common", "normal-specific", "undefined"))) && (all(levels(important_df$Phi_name) == c("common", "normal-specific", "undefined")))){
    personalized_palette = c("#619CFF", "#00BA38", "#808080")
  } else if ((length(levels(important_df$Phi_name)) == length(c("normal-specific"))) && (all(levels(important_df$Phi_name) == c("normal-specific")))){
    personalized_palette = c("#00BA38")
  } else if ((length(levels(important_df$Phi_name)) == length(c("common", "disease-specific", "normal-specific", "undefined"))) && (all(levels(important_df$Phi_name) == c("common", "disease-specific", "normal-specific", "undefined")))){
    personalized_palette = c("#619CFF", "#F8766D", "#00BA38", "#808080")
  } else {
    personalized_palette = c("#619CFF", "#F8766D", "#00BA38", "#808080", "#C77CFF")
  }
  # Plot
  important_node_plot = important_df %>%
    ggplot(aes(fill=Phi_name, y=n_evidences, x=size)) + 
      geom_bar(position="fill", stat="identity", width = 0.7, col="black") +
      #geom_bar(position = position_dodge(), stat="identity", col="black", width = 0.6) +
    xlab("Number of samples") +
    ylab("Fraction of evidences") +
    scale_fill_manual(values = personalized_palette) +
    guides(fill=guide_legend(title="Gene category")) +
    theme_linedraw() +
    theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 14, face="bold"), axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title=element_text(size=14, face="bold"))
  plot_file = paste(plots_dir, paste("codina_scipher_ra_gtex_wholeblood_important_", important_node, ".png", sep=""), sep="/")
  ggsave(
    plot_file,
    plot=important_node_plot,
    dpi = 1200,
    width = 9000,
    height = 6000,
    units = c("px")
  )
  print(important_node_plot)
}
```

Calculate the fraction of unchanged / changed genes in each category for each sample size:

```{r}
# Define output table
cols = c("Node", "Phi_name.prev", "size.prev", "Phi_name.curr", "size.curr")
change_of_category_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))
# Remove levels
nodes_filtered_df$size = as.numeric(as.character(nodes_filtered_df$size))
# For each size, compare nodes with the ones of previous size
for (x in seq(sizes_list)){
  size = sizes_list[x]
  nodes_filtered_by_selected_size = nodes_filtered_df %>%
    filter((!(is.na(DiseaseName.no.sp.char))) & (size == !!size)) %>%
    select(Node, Phi_name, size)
  if (x == 1){
    previous_nodes_filtered = data.frame(nodes_filtered_by_selected_size)
  } else {
    nodes_filtered_compared_by_sizes = previous_nodes_filtered %>%
      inner_join(nodes_filtered_by_selected_size, by=("Node")) %>%
      rename(c("Phi_name.prev"="Phi_name.x", "Phi_name.curr"="Phi_name.y", "size.prev"="size.x", "size.curr"="size.y"))
    change_of_category_df = rbind(change_of_category_df, nodes_filtered_compared_by_sizes)
    previous_nodes_filtered = data.frame(nodes_filtered_by_selected_size)
  }
}

# Detect changes
change_of_category_df$transition_type = ifelse(change_of_category_df$Phi_name.prev == change_of_category_df$Phi_name.curr, "stable", "change")
change_of_category_df$transition_name = paste(change_of_category_df$Phi_name.prev, change_of_category_df$Phi_name.curr, sep=" / ")

# Print table
change_of_category_file = paste(tables_dir, "codina_scipher_ra_gtex_wholeblood_changes.txt", sep="/")
change_of_category_df %>% fwrite(change_of_category_file)
head(change_of_category_df)
```

```{r}
total_stable_change_genes_df = change_of_category_df %>%
  group_by(size.prev) %>% 
  mutate(n_total = n()) %>%
  ungroup() %>%
  select(Node, Phi_name.prev, size.prev, transition_type, n_total) %>% 
  group_by(size.prev, transition_type, n_total) %>% 
  summarize(n_genes = n()) %>% 
  ungroup() %>%
  mutate(frac_genes = n_genes / n_total)
total_stable_change_genes_df$Phi_name.prev = "all"

categories_stable_change_genes_df = change_of_category_df %>%
  group_by(size.prev) %>% 
  mutate(n_total = n()) %>%
  ungroup() %>%
  filter(transition_type == "stable") %>%
  select(Node, Phi_name.prev, size.prev, transition_type, n_total) %>% 
  group_by(Phi_name.prev, size.prev, transition_type, n_total) %>% 
  summarize(n_genes = n()) %>% 
  ungroup() %>%
  mutate(frac_genes = n_genes / n_total)

rbind(total_stable_change_genes_df, categories_stable_change_genes_df) %>%
  filter(transition_type == "stable") %>%
  ggplot(aes(x = size.prev, y = frac_genes, col = Phi_name.prev)) + 
  geom_line(aes(size = Phi_name.prev)) +
  xlab("Number of samples") +
  ylab("Fraction of stable genes") +
  scale_size_manual(values = c("all" = 1, "common" = 0.5, "disease-specific" = 0.5, "normal-specific" = 0.5, "undefined" = 0.5)) +
  scale_color_manual(values = c("#CD9600", "#619CFF", "#F8766D", "#00BA38", "#808080")) +
  guides(col=guide_legend(title="Gene category"), size=FALSE) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 14, face="bold"), axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title=element_text(size=14, face="bold"))
plot_file = paste(plots_dir, "codina_scipher_ra_gtex_wholeblood_stable_genes.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```




