---
title: "Test links revealed by probability"
author: "Joaquim Aguirre-Plans"
date: "4/16/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Given a probability to reveal links and a given number of chances to reveal each link (number of samples), how many samples do we need to reveal all links.

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
require(magrittr)
library(tidyr)
library(igraph)
set.seed(1510)
`%ni%` <- Negate(`%in%`)
```

### Define variables

```{r}
plots_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots'
tables_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/tables'

prob_list = c(0.0001, 0.001, 0.005, 0.01, 0.1)
size_list = seq(20, 1000, 20)
links = seq(1, 20000, 1)

# Define palette using the scale_fill_brewer Oranges palette and selecting manually the colors using RColorBrewer
# https://stackoverflow.com/questions/29466239/how-to-set-the-color-range-of-scale-colour-brewer-in-ggplot2-palette-selecte 
blue_palette = RColorBrewer::brewer.pal(n = 9, "Blues") # I get 8 and exclude the 4 first
name2color <- c(
  "1e-04" = blue_palette[2],
  "0.001" = blue_palette[3],
  "0.005" = blue_palette[5],
  "0.01" = blue_palette[7],
  "0.1" = blue_palette[9]
)
```

### Calculate values of the random model

```{r}
cols = c("link", "result", "size", "prob")
links_revealed_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))

for(prob_selected in prob_list){
  for(size_selected in size_list){
    # Function to calculate if a link is revealed according to a probability and a number of times to try. If res is > 0, the link is revealed
    calculate_prob <- function(i) {  # the argument i is not used
      res = sum(sample(c(0,1), size=size_selected, replace=TRUE, prob=c(1-prob_selected, prob_selected)))
    }
    # Use sapply to run calculate_prob for all links and concatenate each result
    links_revealed_df = rbind(links_revealed_df, data.frame(link=links, result=sapply(links,calculate_prob), size=size_selected, prob=prob_selected))
  }
}
```

Count the results:

```{r}
num_links_per_size = links_revealed_df %>%
  filter(result > 0) %>%
  group_by(size, prob) %>%
  summarize(n_links=n()) %>%
  ungroup() %>%
  as.data.frame()

num_links_per_size_file = paste(tables_dir, 
                                "random_model_not_preserving_links.txt", 
                                sep = "/")

num_links_per_size %>% 
  fwrite(num_links_per_size_file)

head(num_links_per_size)
```

Count the results keeping the link if it has already been revealed:

```{r}
cols = c("link", "result", "size", "prob", "link_revealed")
links_revealed_preserving_df = data.frame(matrix(ncol = length(cols), 
                                                 nrow = 0, 
                                                 dimnames = list(NULL, cols)))

for(size_selected in size_list){
  links_revealed_preserving_df = 
    rbind(links_revealed_preserving_df, 
          links_revealed_df %>%
            filter(size <= size_selected) %>%
            group_by(link, prob) %>%
            mutate(sum_result = sum(result)) %>%
            ungroup() %>%
            filter(size == size_selected) %>%
            mutate(link_revealed = if_else(sum_result > 0, 1, 0)) %>%
            select(-sum_result)
          )
}
```

```{r}
num_links_per_size_preserving = links_revealed_preserving_df %>%
  filter(link_revealed > 0) %>%
  group_by(size, prob) %>%
  summarize(n_links=n())

num_links_per_size_preserving_file = paste(tables_dir, 
                                "random_model_preserving_links.txt", 
                                sep = "/")

num_links_per_size_preserving %>% 
  fwrite(num_links_per_size_preserving_file)

head(num_links_per_size_preserving)
```

### Plot Random Model

Links vs. sample size (not preserving links revealed): 

```{r}
# Plot with legend
num_links_per_size = num_links_per_size %>%
  mutate(prob = factor(prob, 
                       levels = as.character(sort(unique(num_links_per_size$prob)))))

link_reveal_plot = num_links_per_size %>%
  ggplot(aes(x = log10(size), y = n_links, col = prob)) +
  #geom_point(alpha=0.8, size=3) +
  geom_line(size = 1) +
  scale_color_manual(name = "Prob. reveal link", 
                     values = as.vector(name2color[levels(factor(num_links_per_size$prob))])) +
  theme_linedraw() +
  #ggtitle("Links revealed vs. sample size") +
  xlab("Num. samples (Log10)") +
  ylab("Num. links revealed") +
  theme(aspect.ratio = 1, 
        plot.title =  element_text(size = 20, face="bold"), 
        axis.title = element_text(size = 17, face="bold"), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        legend.title=element_text(size=16, face="bold"),
        panel.grid.major=element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(family = "Helvetica")
        )

plot_file = paste(plots_dir, 
                  "random_model_not_preserving_links.png", 
                  sep="/")
ggsave(
  plot_file,
  plot = link_reveal_plot,
  dpi = 1200,
  width = 10000,
  height = 6000,
  units = c("px")
)
print(link_reveal_plot)

```

Links vs. sample size (preserving links revealed): 

```{r}
# Plot with legend
num_links_per_size_preserving = num_links_per_size_preserving %>%
  mutate(prob = factor(prob, 
                       levels = as.character(sort(unique(num_links_per_size_preserving$prob)))))

link_reveal_plot = num_links_per_size_preserving %>%
  ggplot(aes(x=size, y=log10(n_links), col=prob)) +
  #geom_point(alpha=0.8, size=3) +
  geom_line(size = 1) +
  scale_color_manual(name = "Prob. reveal link", 
                     values = as.vector(name2color[levels(factor(num_links_per_size_preserving$prob))])) +
  theme_linedraw() +
  #ggtitle("Links revealed vs. sample size (preserving links)") +
  xlab("Num. samples (Log10)") +
  ylab("Num. links revealed") +
  theme(aspect.ratio = 1, 
        plot.title =  element_text(size = 20, face="bold"), 
        axis.title = element_text(size = 17, face="bold"), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        legend.title=element_text(size=16, face="bold"),
        panel.grid.major=element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(family = "Helvetica")
        )

plot_file = paste(plots_dir, 
                  "random_model_preserving_links.png", 
                  sep="/")
ggsave(
  plot_file,
  plot = link_reveal_plot,
  dpi = 1200,
  width = 10000,
  height = 6000,
  units = c("px")
)
print(link_reveal_plot)
```

Links vs. sample size (combining revealing links and not revealing links): 

```{r}
# Plot with legend
num_links_per_size_comb = 
  rbind((num_links_per_size %>% 
          mutate(preserve_links = TRUE)),
        (num_links_per_size_preserving %>%
          mutate(preserve_links = FALSE) %>%
          as.data.frame())
        )

link_reveal_plot = num_links_per_size_comb %>%
  ggplot(aes(x = log10(size), y = n_links, col = prob, linetype = preserve_links)) +
  #geom_point(alpha=0.8, size=3) +
  geom_line(size = 1) +
  scale_color_manual(name = "Prob. reveal link", 
                     values = as.vector(name2color[levels(factor(num_links_per_size_preserving$prob))])) +
  theme_linedraw() +
  #ggtitle("Links revealed vs. sample size (preserving links)") +
  xlab("Num. samples (Log10)") +
  ylab("Num. links revealed") +
  guides(linetype = guide_legend(title = "Preserve links")) +
  theme(aspect.ratio = 1, 
        plot.title =  element_text(size = 20, face="bold"), 
        axis.title = element_text(size = 17, face="bold"), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        legend.title=element_text(size=16, face="bold"),
        panel.grid.major=element_blank(), 
        panel.grid.minor = element_blank(),
        text = element_text(family = "Helvetica")
        )

plot_file = paste(plots_dir, 
                  "random_model.png", 
                  sep="/")
ggsave(
  plot_file,
  plot = link_reveal_plot,
  dpi = 1200,
  width = 10000,
  height = 6000,
  units = c("px")
)
print(link_reveal_plot)
```