---
title: "Create Figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Create figures for the manuscript, extended abstract and poster.

```{r}
library(data.table)
library(dplyr)
library(igraph)
library(ggplot2)
require(ggrepel)
require(magrittr)
library(tidyr)
set.seed(1510)
options(bitmapType='cairo')
`%ni%` <- Negate(`%in%`)
```

### Load data

Un-normalized data:

```{r}
input_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/scripts/SampleSizeShiny/data'
plots_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/plots'
numbers_file = paste(input_dir, 'dataset_numbers_complete_graph.txt', sep='/')
numbers_df = fread(numbers_file)
numbers_df$dataset = tolower(numbers_df$dataset)

input_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/scripts/SampleSizeShiny/data/example_pearson_pval_0.05'
topology_results_file = paste(input_dir, 'topology_results_pearson_pval_0.05.txt', sep='/')
results_selected_df = fread(topology_results_file)
topology_results_by_size_file = paste(input_dir, 'topology_results_mean_pearson_pval_0.05.txt', sep='/')
topology_results_selected_by_size_df = fread(topology_results_by_size_file)
predictions_file = paste(input_dir, 'predictions_pearson_pval_0.05.txt', sep='/')
predicted_results_df = fread(predictions_file)
analytical_results_file = paste(input_dir, 'analytical_model_results_pearson_pval_0.05.txt', sep='/')
topology_results_selected_analytical_df = fread(analytical_results_file)
analytical_summary_file = paste(input_dir, 'analytical_model_summary_pearson_pval_0.05.txt', sep='/')
analytical_model_summary_df = fread(analytical_summary_file)
analytical_regression_results_file = paste(input_dir, 'analytical_model_regression_results_pearson_pval_0.05.txt', sep='/')
stretched_exponential_regression_df = fread(analytical_regression_results_file)
theoretical_sample_size_file = paste(input_dir, 'theoretical_sample_size_for_correlations_of_datasets.txt', sep='/')
sample_size_correlation_df = fread(theoretical_sample_size_file)
```

Normalized data:

```{r}
topology_type_normalization = "divide.L"
input_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/scripts/SampleSizeShiny/data/example_pearson_pval_0.05'
topology_results_file = paste(input_dir, '/', 'topology_results_norm_', topology_type_normalization, '_pearson_pval_0.05.txt', sep='')
results_selected_norm_df = fread(topology_results_file)
topology_results_by_size_file = paste(input_dir, '/', 'topology_results_mean_norm_', topology_type_normalization, '_pearson_pval_0.05.txt', sep='')
topology_results_selected_by_size_norm_df = fread(topology_results_by_size_file)
predictions_file = paste(input_dir, '/', 'predictions_', topology_type_normalization, '_pearson_pval_0.05.txt', sep='')
predicted_results_norm_df = fread(predictions_file)
analytical_results_file = paste(input_dir, '/', 'analytical_model_results_', topology_type_normalization, '_pearson_pval_0.05.txt', sep='')
topology_results_selected_analytical_norm_df = fread(analytical_results_file)
```

Number of networks and maximum sample size:

```{r}
nrow((results_selected_df %>% dplyr::select(method, type_dataset, size, rep) %>% unique()))
max(((results_selected_df %>% dplyr::select(method, type_dataset, size, rep) %>% unique()))$size)
```

Number of networks and maximum sample size without counting TCGA full dataset:

```{r}
nrow((results_selected_df %>% dplyr::select(method, type_dataset, size, rep) %>% filter(!(type_dataset == "tcga:tcga")) %>% unique()))
max(((results_selected_df %>% dplyr::select(method, type_dataset, size, rep) %>% filter(!(type_dataset == "tcga:tcga")) %>% unique()))$size)
```

### Make plots

#### Curve plot

```{r}
#datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca")
datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca", "scipher:complete.dataset")
results_selected_norm_df %>%
  filter(type_dataset %in% datasets_selected) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "R. arthritis")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:whole.blood", "GTEx: Whole blood")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:artery.tibial", "GTEx: Artery tibial")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-coad", "TCGA: Colon cancer")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-brca", "TCGA: Breast cancer")) %>%
  ggplot(aes(x=size, y=unnorm, col=type_dataset)) +
  geom_point(alpha=0.5, size=3) +
  #scale_x_continuous(trans = scales::log10_trans()) +
  scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  theme_linedraw() +
  xlab("Num. samples") +
  #xlab("log(N)") +
  ylab("Num. significant correlations") +
  guides(col=guide_legend(title="Dataset")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))
plot_file = paste(plots_dir, "curve_pearson_pval_0.05.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 10000,
  height = 6000,
  units = c("px")
)
```

#### Regression plot

```{r}
#datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca")
datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca", "scipher:complete.dataset")
#model_selected = "Stretched exponential (by optimization)"
model_selected = "Stretched exponential (by linear fit)"
stretched_exponential_regression_df %>% inner_join((topology_results_selected_analytical_df %>% dplyr::select(type_dataset, model, a, b) %>% unique()), by=c("type_dataset", "model")) %>%
  filter((type_dataset %in% datasets_selected) & (model == model_selected)) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "R. arthritis")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:whole.blood", "GTEx: Whole blood")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:artery.tibial", "GTEx: Artery tibial")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-coad", "TCGA: Colon cancer")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-brca", "TCGA: Breast cancer")) %>%
  ggplot(aes(x=x, y=y, col=type_dataset)) +
  geom_point(alpha=0.5, size=3) +
  geom_abline(aes(intercept=b, slope=a, col=type_dataset), size=1) +
  theme_linedraw() +
  xlab("log(N)") +
  ylab("log[ log(L) - log(S) ]") +
  scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  guides(col=guide_legend(title="Dataset")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))
plot_file = paste(plots_dir, "regression_pearson_pval_0.05.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9500,
  height = 6000,
  units = c("px")
)
```

```{r}
#datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca")
datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca", "scipher:complete.dataset")
#model_selected = "Stretched exponential (by optimization)"
model_selected = "Stretched exponential (by linear fit)"
analytical_model_summary_df$L_vs_total = abs((analytical_model_summary_df$unnorm_L - analytical_model_summary_df$total_num_edges)) / analytical_model_summary_df$total_num_edges
analytical_model_summary_df %>% 
  dplyr::select(type_dataset, model, a, b, L, L_vs_total, adj.r.squared, relative.error.mean) %>% 
  filter((type_dataset %in% datasets_selected) & (model == model_selected)) %>% unique()
```

#### Prediction plot

```{r}
#datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca")
datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca", "scipher:complete.dataset")
#model_selected = "Stretched exponential (by optimization)"
model_selected = "Stretched exponential (by linear fit)"
results_selected_norm_df %>% right_join((predicted_results_norm_df %>% dplyr::select(type_dataset, model, model_result, size) %>% unique()), by=c("type_dataset", "size")) %>%
  filter((type_dataset %in% datasets_selected) & (model == model_selected)) %>%
#results_selected_norm_df %>%
#  filter(type_dataset %in% datasets_selected) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "R. arthritis")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:whole.blood", "GTEx: Whole blood")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:artery.tibial", "GTEx: Artery tibial")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-coad", "TCGA: Colon cancer")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-brca", "TCGA: Breast cancer")) %>%
  ggplot(aes(x=size, y=num_edges, col=type_dataset)) +
  geom_point(alpha=0.5, size=3) +
  #geom_line(data=(predicted_results_norm_df %>% filter((type_dataset %in% datasets_selected) & (model == model_selected)) %>% unique()), aes(x=size, y=model_result, col=type_dataset), size=1) +
  geom_line(aes(x=size, y=model_result, col=type_dataset), size=1) +
  scale_x_continuous(trans = scales::log10_trans()) +
  scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  theme_linedraw() +
  xlab("Num. samples (Log)") +
  ylab("Fraction of significant correlations") +
  guides(col=guide_legend(title="Dataset")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))
plot_file = paste(plots_dir, "prediction_pearson_pval_0.05.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9500,
  height = 6000,
  units = c("px")
)
```

#### Comparison of models: stretched exponential vs. logarithm

```{r}
models_selected = c("Stretched exponential (by linear fit)", "Logarithmic", "Mean")
dataset_selected = "gtex:whole.blood"

# Join predicted results with mean
predicted_results_norm_and_mean_df = rbind((topology_results_selected_by_size_norm_df %>% dplyr::select(type_dataset, size, mean_norm) %>% unique() %>% rename("model_result" = "mean_norm") %>% mutate(model="Mean")), (predicted_results_norm_df %>% dplyr::select(type_dataset, size, model_result, model) %>% unique())) %>% 
  filter(model %in% models_selected) %>%
  mutate(model = replace(model, model == "Stretched exponential (by linear fit)", "Stretched exp.")) %>%
  mutate(model = replace(model, model == "Logarithmic", "Logarithm"))
predicted_results_norm_and_mean_df$model <- factor(predicted_results_norm_and_mean_df$model, levels = c("Mean", "Stretched exp.", "Logarithm"))

# Process and plot
results_selected_norm_df %>% 
  right_join(predicted_results_norm_and_mean_df, by=c("type_dataset", "size")) %>%
  filter(type_dataset == dataset_selected) %>%
  filter(size <= max((results_selected_norm_df %>% filter(type_dataset == dataset_selected))$size)) %>%
  filter(size >= min((results_selected_norm_df %>% filter(type_dataset == dataset_selected))$size)) %>%
  ggplot(aes(x=size, y=num_edges, col=model)) +
  geom_point(alpha=0.5, size=3, col="#91faad") +
  geom_line(aes(x=size, y=model_result, col=model), size=1) +
  scale_x_continuous(trans = scales::log10_trans()) +
  scale_color_manual(values = c("#00f549", "#02b0f5", "#fc9403")) +
  theme_linedraw() +
  xlab("Num. samples (Log)") +
  ylab("Fraction of significant correlations") +
  #guides(col=guide_legend(title="Model")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_blank(), legend.position="bottom")
plot_file = paste(plots_dir, "comparison_models_pearson_pval_0.05.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 6600,
  height = 6700,
  units = c("px")
)

```

#### a vs. fraction of significant correlations plot

```{r}
#'  calculate_predictions_using_stretched_exponential_model_optimized
#'  Formula to calculate exponential decay of significant interactions from a list of sample sizes.
#'  This formula requires the use of the L parameter
#'  @param x List of sample sizes.
#'  @param a Slope coefficient.
#'  @param b Intercept coefficient.
#'  
calculate_predictions_using_stretched_exponential_model_optimized = function(x, L, a, b){
  y = exp(log(L) - exp(b+a*log(x)))
  return(y)
}

#'  calculate_prediction_from_analytical_model
#'  Function to calculate predictions using a specific analytical model
#'  @param model Name of the model used.
#'  @param x_list List of values in the x axis (e.g. size)
#'  @param a Parameter a.
#'  @param b Parameter b.
#'  @param L Parameter L.
#'  
calculate_prediction_from_analytical_model = function(model, x_list, a, b, L){
  if(model == "Logarithmic"){
    prediction_result = (log(x_list)*a + b)
  } else if((model == "Stretched exponential (by optimization)") | (model == "Stretched exponential (by linear fit)")){
    prediction_result = calculate_predictions_using_stretched_exponential_model_optimized(x=x_list, L=L, a=a, b=b)
  } else if(model == "Stretched exponential (without L)"){
    prediction_result = calculate_predictions_using_stretched_exponential_model_without_L(x=x_list, a=a, b=b)
  }
  return(prediction_result)
}
```

```{r}
#datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca")
datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca", "scipher:complete.dataset")
#model_selected = "Stretched exponential (by optimization)"
model_selected = "Stretched exponential (by linear fit)"
sample_size_vs_a_df = analytical_model_summary_df %>% inner_join(sample_size_correlation_df, by=c("type_dataset"="dataset")) %>% 
  #filter((type_dataset %in% datasets_selected) & (model == model_selected)) 
  filter(model == model_selected)

# Calculate number of edges for each critical correlation using model
sample_size_vs_a_df$num_edges_from_statistical_corrected = calculate_prediction_from_analytical_model(model=model_selected, x_list=sample_size_vs_a_df$sample_size_statistical_corrected, a=sample_size_vs_a_df$a, b=sample_size_vs_a_df$b, L=sample_size_vs_a_df$L)
sample_size_vs_a_df$num_edges_from_statistical_corrected = sample_size_vs_a_df$num_edges_from_statistical_corrected * sample_size_vs_a_df$max_value_in_dataset
sample_size_vs_a_df$num_edges_from_statistical_corrected_norm = sample_size_vs_a_df$num_edges_from_statistical_corrected / sample_size_vs_a_df$total_num_edges
```

```{r}
type_correlation_selected = "weak"
#sample_size_vs_a_df$data_col = ifelse(sample_size_vs_a_df$type_dataset == "gtex:whole.blood", "#E69F00", ifelse(sample_size_vs_a_df$type_dataset == "gtex:artery.tibial", "#D55E00", ifelse(sample_size_vs_a_df$type_dataset == "tcga:tcga-coad", "#56B4E9", ifelse(sample_size_vs_a_df$type_dataset == "tcga:tcga-brca", "#0072B2", "black"))))

sample_size_vs_a_df %>% 
  filter(type_correlation == type_correlation_selected) %>% 
  filter(!(type_dataset == "tcga:tcga")) %>% 
  mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "scipher")) %>%
  ggplot(aes(x=a, y=num_edges_from_statistical_corrected_norm)) + 
  geom_point() +
  geom_point(data = . %>% filter(type_dataset == "tcga:tcga-coad"), color = "#56B4E9") +
  geom_point(data = . %>% filter(type_dataset == "tcga:tcga-brca"), color = "#0072B2") +
  geom_point(data = . %>% filter(type_dataset == "scipher"), color = "#44AA99") +
  geom_point(data = . %>% filter(type_dataset == "gtex:whole.blood"), color = "#E69F00") +
  geom_point(data = . %>% filter(type_dataset == "gtex:artery.tibial"), color = "#D55E00") +
  xlab(expression(alpha)) +
  ylab("Fraction of correlations above 0.2") +
  #guides(col=guide_legend(title="Dataset")) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold")) +
  geom_text_repel(
    data = subset((sample_size_vs_a_df %>% filter(type_correlation == type_correlation_selected)), type_dataset == "gtex:whole.blood"),
    aes(label = type_dataset), col = "#E69F00",
    size = 5,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.3, "lines")
  ) +
  geom_text_repel(
    data = subset((sample_size_vs_a_df %>% filter(type_correlation == type_correlation_selected)), type_dataset == "gtex:artery.tibial"),
    aes(label = type_dataset), col = "#D55E00",
    size = 5,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  ) +
  geom_text_repel(
    data = subset((sample_size_vs_a_df %>% filter(type_correlation == type_correlation_selected) %>% mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "scipher"))), type_dataset == "scipher"),
    aes(label = type_dataset), col = "#44AA99",
    size = 5,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.3, "lines")
  ) +
  geom_text_repel(
    data = subset((sample_size_vs_a_df %>% filter(type_correlation == type_correlation_selected)), type_dataset == "tcga:tcga-brca"),
    aes(label = type_dataset), col = "#0072B2",
    size = 5,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  ) +
  geom_text_repel(
    data = subset((sample_size_vs_a_df %>% filter(type_correlation == type_correlation_selected)), type_dataset == "tcga:tcga-coad"),
    aes(label = type_dataset), col = "#56B4E9",
    size = 5,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.3, "lines")
  )

plot_file = paste(plots_dir, "a_vs_fraction_sig_correlations_pearson_pval_0.05.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 6400,
  height = 6000,
  units = c("px")
)
```

```{r}
type_correlation_selected = "weak"
#sample_size_vs_a_df$data_col = ifelse(sample_size_vs_a_df$type_dataset == "gtex:whole.blood", "#E69F00", ifelse(sample_size_vs_a_df$type_dataset == "gtex:artery.tibial", "#D55E00", ifelse(sample_size_vs_a_df$type_dataset == "tcga:tcga-coad", "#56B4E9", ifelse(sample_size_vs_a_df$type_dataset == "tcga:tcga-brca", "#0072B2", "black"))))

sample_size_vs_a_df %>% 
  filter(type_correlation == type_correlation_selected) %>% 
  filter(!(type_dataset == "tcga:tcga")) %>% 
  mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "R. arthritis")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:whole.blood", "GTEx: Whole blood")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:artery.tibial", "GTEx: Artery tibial")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-coad", "TCGA: Colon cancer")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-brca", "TCGA: Breast cancer")) %>%
  ggplot(aes(x=a, y=num_edges_from_statistical_corrected_norm)) + 
  geom_point() +
  geom_point(data = . %>% filter(type_dataset == "TCGA: Colon cancer"), color = "#56B4E9") +
  geom_point(data = . %>% filter(type_dataset == "TCGA: Breast cancer"), color = "#0072B2") +
  geom_point(data = . %>% filter(type_dataset == "R. arthritis"), color = "#44AA99") +
  geom_point(data = . %>% filter(type_dataset == "GTEx: Whole blood"), color = "#E69F00") +
  geom_point(data = . %>% filter(type_dataset == "GTEx: Artery tibial"), color = "#D55E00") +
  xlab(expression(alpha)) +
  ylab("Fraction of correlations above 0.2") +
  #guides(col=guide_legend(title="Dataset")) +
  theme_linedraw() +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))

plot_file = paste(plots_dir, "a_vs_fraction_sig_correlations_pearson_pval_0.05_no_labels.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 6400,
  height = 6000,
  units = c("px")
)
```

#### Levels of correlations plot

Define function to calculate critical sample size for correlation:

```{r}
calculate_sample_size_for_pearson_correlation_using_t_distribution_with_optimization = function(r, total_num_edges, alpha=0.05, N_guess=c(10000)){
  optimize_N = function(par){
    N = par[1]
    t_guess = qt(alpha, df=N-2, lower.tail = F)
    t = r * sqrt((N-2)) / sqrt((1-r**2))
    return((abs(t-t_guess)))
  }
  res = optim(par=N_guess, fn=optimize_N, method="Brent", lower=3, upper=50000)
  return(res$par)
}

calculate_predictions_using_stretched_exponential_model_optimized = function(x, L, a, b){
  y = exp(log(L) - exp(b+a*log(x)))
  return(y)
}
```

Calculate sample size for different levels of correlation and different datasets:

```{r}
type_correlation_df = data.frame(type_correlation=c("weak", "moderate", "strong", "very strong"), lower_val=c(0.2,0.4,0.6,0.8), upper_val=c(0.4,0.6,0.8,NA))
types_correlation = c("weak", "moderate", "strong", "very strong")

cols = c("dataset", "type_correlation", "correlation", "sample_size_statistical", "sample_size_statistical_corrected")
sample_size_correlation_df = data.frame(matrix(ncol=length(cols),nrow=0, dimnames=list(NULL, cols)))
for (dataset in numbers_df$dataset){
  total_num_edges = (numbers_df %>% filter(dataset == !!dataset))$total_num_edges
  for (type_correlation in types_correlation){
    r = (type_correlation_df %>% filter(type_correlation == !!type_correlation))$lower_val
    N = calculate_sample_size_for_pearson_correlation_using_t_distribution_with_optimization(r=r, total_num_edges=total_num_edges, alpha=0.05, N_guess=c(10000))
    N_corrected = calculate_sample_size_for_pearson_correlation_using_t_distribution_with_optimization(r=r, total_num_edges=total_num_edges, alpha=0.05/total_num_edges, N_guess=c(10000))
    sample_size_correlation_df <- rbind(sample_size_correlation_df, data.frame(dataset=dataset, type_correlation=type_correlation, correlation=r, sample_size_statistical=N, sample_size_statistical_corrected=N_corrected))
  }
}

print(sample_size_correlation_df)
```

```{r}
dataset_selected = "gtex:whole.blood"
model_selected = "Stretched exponential (by linear fit)"
results_analytical_gtex_wb = topology_results_selected_analytical_norm_df %>% 
  filter((model == model_selected) & (type_dataset == dataset_selected)) %>%
  select(model, type_dataset, a, b, L, max_value_in_dataset) %>% unique()
sample_size_correlation_gtex_wb = sample_size_correlation_df %>%
  filter((dataset == dataset_selected) & (!(type_correlation == "weak")))

predictions_convergence_gtex_wb_norm = calculate_predictions_using_stretched_exponential_model_optimized(x=sample_size_correlation_gtex_wb$sample_size_statistical_corrected, L=results_analytical_gtex_wb$L, a=results_analytical_gtex_wb$a, b=results_analytical_gtex_wb$b)
predictions_convergence_gtex_wb_df = data.frame(size=round(sample_size_correlation_gtex_wb$sample_size_statistical_corrected), num_edges=predictions_convergence_gtex_wb_norm * results_analytical_gtex_wb$max_value_in_dataset, num_edges_norm=predictions_convergence_gtex_wb_norm)
predictions_convergence_gtex_wb_df$num_edges_norm = (predictions_convergence_gtex_wb_df$num_edges/results_analytical_gtex_wb$max_value_in_dataset)/results_analytical_gtex_wb$L
predictions_convergence_gtex_wb_df

predicted_results_wb_df = predicted_results_df %>% filter((model == model_selected) & (type_dataset == dataset_selected))
predicted_results_wb_df$model_result_norm = (predicted_results_wb_df$model_result/results_analytical_gtex_wb$max_value_in_dataset)/results_analytical_gtex_wb$L
```

```{r}
dataset_selected = "Whole.Blood"
all_topology_results_file = '/home/j.aguirreplans/Projects/Scipher/SampleSize/scripts/SampleSizeShiny/data/analysis_topology.csv'
threshold_selected = 0.05

all_results_df = fread(all_topology_results_file)
results_gtex_wb = all_results_df %>% 
  filter((type_dataset == dataset_selected) & (threshold == threshold_selected) & (type_correlation %in% c("weak-moderate-strong-very strong", "moderate-strong-very strong", "strong-very strong", "very strong")))
results_gtex_wb$num_edges_norm = (results_gtex_wb$num_edges/results_analytical_gtex_wb$max_value_in_dataset)/results_analytical_gtex_wb$L

results_gtex_wb %>% 
  mutate(type_correlation = replace(type_correlation, type_correlation == "weak-moderate-strong-very strong", "\u2265 0.2")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "moderate-strong-very strong", "\u2265 0.4")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "strong-very strong", "\u2265 0.6")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "very strong", "\u2265 0.8")) %>%
  ggplot(aes(x=size, y=num_edges_norm, col=type_correlation)) +
  geom_point(alpha=0.5, size=3) +
  geom_line(data=(predicted_results_wb_df %>% filter(size <= max((all_results_df %>% filter((type_dataset == dataset_selected) & (threshold == threshold_selected)))$size))), aes(x=size, y=model_result_norm), size=1, color='red') +
  geom_point(data=predictions_convergence_gtex_wb_df, 
           aes(x=size,y=num_edges_norm), 
           color='red',
           size=3) +
  #scale_x_continuous(trans = scales::log10_trans()) +
  #scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  theme_linedraw() +
  xlab("Number of samples") +
  #xlab("log(N)") +
  ylab("Fraction of significant correlations") +
  guides(col=guide_legend(title="Correlation level")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))

plot_file = paste(plots_dir, "correlation_levels_pearson_pval_0.05_data_gtexwholeblood.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```


#### Disease modules plot

```{r}
all_disease_genes_results_file = '/home/j.aguirreplans/Projects/Scipher/SampleSize/scripts/SampleSizeShiny/data/analysis_disease_genes.csv'
threshold_selected = 0.05
dataset_selected = "scipher:complete.dataset"

all_disease_genes_results_df = fread(all_disease_genes_results_file)
all_disease_genes_results_df$type_dataset = paste(all_disease_genes_results_df$dataset, all_disease_genes_results_df$type_dataset, sep=":") # Join dataset and type_dataset
all_disease_genes_results_df$type_dataset = tolower(all_disease_genes_results_df$type_dataset)

results_scipher_df = all_disease_genes_results_df %>%
  filter((type_dataset == dataset_selected) & (threshold == threshold_selected))
head(results_scipher_df)
```

```{r}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
results_scipher_df %>%
  filter(type_correlation == "moderate-strong-very strong") %>%
  group_by(size, disease) %>%
  summarize(mean_pval = mean(disease_lcc_pvalue)) %>%
  ggplot(aes(x=size, y=abs(log10(mean_pval)), col=disease)) +
  #geom_point(alpha=0.5, size=3) +
  geom_line(size=1) +
  geom_hline(yintercept=as.numeric(abs(log10(0.05))), linetype="dashed", color="red", size=0.5) +
  #scale_x_continuous(trans = scales::log10_trans()) +
  scale_color_manual(values = cbbPalette) +
  theme_linedraw() +
  xlab("Number of samples") +
  #xlab("log(N)") +
  ylab("Disease LCC log p-value (abs)") +
  guides(col=guide_legend(title="Disease")) +
  theme(plot.title =  element_text(size = 17, face="bold"), 
        axis.title = element_text(size = 16, face="bold"), 
        axis.text = element_text(size = 15), 
        #legend.position = "bottom",
        legend.text = element_text(size = 14), 
        legend.title=element_text(size=15, face="bold"))
plot_file = paste(plots_dir, "disease_module_significance_by_diseases_pearson_pval_0.05_data_sciphercomplete.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9700,
  height = 6000,
  units = c("px")
)
```

```{r}
results_scipher_df %>%
  filter(disease == "arthritis rheumatoid") %>%
  filter((type_dataset == dataset_selected) & (threshold == threshold_selected) & (type_correlation %in% c("weak-moderate-strong-very strong", "moderate-strong-very strong", "strong-very strong", "very strong"))) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "weak-moderate-strong-very strong", "\u2265 0.2")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "moderate-strong-very strong", "\u2265 0.4")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "strong-very strong", "\u2265 0.6")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "very strong", "\u2265 0.8")) %>%
  group_by(size, type_correlation) %>%
  mutate(mean_pval = mean(disease_lcc_pvalue)) %>%
  ungroup() %>%
  ggplot() +
  geom_point(aes(x=size, y=abs(log10(disease_lcc_pvalue)), col=type_correlation), alpha=0.4, size=3) +
  geom_line(aes(x=size, y=abs(log10(mean_pval)), col=type_correlation), size=1) +
  geom_hline(yintercept=as.numeric(abs(log10(0.05))), linetype="dashed", color="red", size=0.5) +
  #scale_x_continuous(trans = scales::log10_trans()) +
  #scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  theme_linedraw() +
  xlab("Number of samples") +
  #xlab("log(N)") +
  ylab("Disease LCC log p-value (abs)") +
  guides(col=guide_legend(title="Correlation level")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))
plot_file = paste(plots_dir, "disease_module_significance_by_correlations_pearson_pval_0.05_data_sciphercomplete.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9000,
  height = 6000,
  units = c("px")
)
```

```{r}
results_scipher_df %>%
  filter(disease == "arthritis rheumatoid") %>%
  filter((type_dataset == dataset_selected) & (threshold == threshold_selected) & (type_correlation %in% c("weak-moderate-strong-very strong", "moderate-strong-very strong", "strong-very strong", "very strong"))) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "weak-moderate-strong-very strong", "\u2265 0.2")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "moderate-strong-very strong", "\u2265 0.4")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "strong-very strong", "\u2265 0.6")) %>%
  mutate(type_correlation = replace(type_correlation, type_correlation == "very strong", "\u2265 0.8")) %>%
  group_by(size, type_correlation) %>%
  mutate(mean_rlcc = mean(fraction_disease_lcc_nodes)) %>%
  ungroup() %>%
  ggplot() +
  geom_point(aes(x=size, y=fraction_disease_lcc_nodes, col=type_correlation), alpha=0.4, size=3) +
  geom_line(aes(x=size, y=mean_rlcc, col=type_correlation), size=1) +
  #scale_x_continuous(trans = scales::log10_trans()) +
  #scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  theme_linedraw() +
  xlab("Number of samples") +
  #xlab("log(N)") +
  ylab("Fraction of disease genes in the LCC") +
  guides(col=guide_legend(title="Correlation level")) +
  theme(plot.title =  element_text(size = 17, face="bold"), axis.title = element_text(size = 16, face="bold"), axis.text = element_text(size = 15), legend.text = element_text(size = 14), legend.title=element_text(size=15, face="bold"))
plot_file = paste(plots_dir, "disease_module_rlcc_by_correlations_pearson_pval_0.05_data_sciphercomplete.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9300,
  height = 6000,
  units = c("px")
)
```

```{r}
datasets_selected = c("gtex:whole.blood", "gtex:artery.tibial", "tcga:tcga-coad", "tcga:tcga-brca", "scipher:complete.dataset")
all_disease_genes_results_df %>%
  filter((disease == "arthritis rheumatoid") & (type_dataset %in% datasets_selected) & (threshold == threshold_selected) & (type_correlation == "moderate-strong-very strong")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "scipher:complete.dataset", "R. arthritis")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:whole.blood", "GTEx: Whole blood")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "gtex:artery.tibial", "GTEx: Artery tibial")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-coad", "TCGA: Colon cancer")) %>%
  mutate(type_dataset = replace(type_dataset, type_dataset == "tcga:tcga-brca", "TCGA: Breast cancer")) %>%
  group_by(size, type_dataset) %>%
  mutate(mean_pval = mean(disease_lcc_pvalue)) %>%
  ungroup() %>%
  ggplot() +
  geom_point(aes(x=size, y=abs(log10(disease_lcc_pvalue)), col=type_dataset), alpha=0.5, size=3) +
  #geom_line(aes(x=size, y=abs(log10(mean_pval)), col=type_dataset), size=1) +
  geom_hline(yintercept=as.numeric(abs(log10(0.05))), linetype="dashed", color="red", size=0.5) +
  #scale_x_continuous(trans = scales::log10_trans()) +
  scale_color_manual(values = c("#D55E00", "#E69F00", "#44AA99", "#0072B2", "#56B4E9")) +
  theme_linedraw() +
  xlab("Number of samples") +
  #xlab("log(N)") +
  ylab("Disease LCC log p-value (abs)") +
  guides(col=guide_legend(title="Dataset")) +
  theme(plot.title =  element_text(size = 17, face="bold"), 
        axis.title = element_text(size = 16, face="bold"), 
        axis.text = element_text(size = 15), 
        legend.text = element_text(size = 14), 
        legend.title=element_text(size=15, face="bold"))

plot_file = paste(plots_dir, "disease_module_significance_by_dataset_pearson_pval_0.05.png", sep="/")
ggsave(
  plot_file,
  dpi = 1200,
  width = 9200,
  height = 6000,
  units = c("px")
)

```


