---
title: "Get targets of a drug indicated for a disease"
author: "Joaquim Aguirre-Plans"
date: '2022-12-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Create the drug-target interactions file of drugs indicated for a specific disease

```{r}
library(data.table)
library(dplyr)
require(magrittr)
library(tidyr)
set.seed(1510)
options(bitmapType='cairo')
`%ni%` <- Negate(`%in%`)
```

### Read files

Read drug indications file (OTP):

```{r}
otp_indications_file = "/work/ccnr/j.aguirreplans/Databases/Open_Targets_Platform/data/out/drug_disease_v2.csv"
otp_indications_df = fread(otp_indications_file) %>%
  rename("chembl_id"="chemId", "chembl_name"="chemName", "efo_id"="efoId", "efo_term"="efoName", "max_phase_indication" = "maxPhaseForIndication") %>%
  mutate(chembl_id=toupper(chembl_id), chembl_name=tolower(chembl_name), efo_id=toupper(efo_id), efo_term=tolower(efo_term)) %>%
  unique()
glimpse(otp_indications_df)
```

Read drug indications file (ChEMBL):

```{r}
chembl_indications_file = "/work/ccnr/j.aguirreplans/Databases/ChEMBL/data/raw/chembl_indications.csv"
chembl_indications_df = fread(chembl_indications_file) %>% 
  rename("chembl_id" = "Parent Molecule ChEMBL ID", "chembl_name" = "Parent Molecule Name", "molecule_type" = "Parent Molecule Type", "max_phase_indication" = "Max Phase for Indication", "first_approval" = "First Approval", "mesh_id" = "MESH ID", "mesh_heading" = "MESH Heading", "efo_id" = "EFO IDs", "efo_term" = "EFO Terms") %>% 
  select(chembl_id, chembl_name, molecule_type, max_phase_indication, first_approval, mesh_id, mesh_heading, efo_id, efo_term) %>%
  mutate(chembl_id=toupper(chembl_id), chembl_name=tolower(chembl_name), molecule_type=tolower(molecule_type), mesh_id=toupper(mesh_id), mesh_heading=tolower(mesh_heading), efo_term=tolower(efo_term)) %>%
  unique()
glimpse(chembl_indications_df)
```

Read drug targets file (DrugBank):

```{r}
drugbank_drug_dictionary_file = "/work/ccnr/j.aguirreplans/Databases/DrugBank/data/out/DB_Drug_Dictionary_v5.1.9.csv"
drugbank_drug_targets_file = "/work/ccnr/j.aguirreplans/Databases/DrugBank/data/out/DB_Drug_Targets_v5.1.9.csv"
drugbank_drug_targets_df = fread(drugbank_drug_targets_file) %>% 
  select(DB_ID, Name, ATC, Status, Gene_Target, P_id, Gene_Desc, Organism, Type, Known_Action, Mech_Of_Action, Indication) %>% 
  rename("DB_Name"="Name") %>%
  mutate(DB_ID=toupper(DB_ID), DB_Name=tolower(DB_Name), ATC=toupper(ATC), Status=tolower(Status), Gene_Target=toupper(Gene_Target), Gene_Desc=tolower(Gene_Desc), Organism=tolower(Organism), Type=tolower(Type), Known_Action=tolower(Known_Action), Mech_Of_Action=tolower(Mech_Of_Action)) %>%
  unique()
glimpse(drugbank_drug_targets_df)
```

And drug synonyms file (DrugBank):

```{r}
drugbank_drug_external_file = "/work/ccnr/j.aguirreplans/Databases/DrugBank/data/out/DB_Drug_External_ID_v5.1.9.csv"
drugbank_drug_external_df = fread(drugbank_drug_external_file, header = TRUE) %>% 
  select(-V1, -Name) %>%
  rename("DB_ID" = "ID", "external_id_source" = "Source", "external_id" = "ID_external") %>%
  mutate(DB_ID=toupper(DB_ID), external_id_source=tolower(external_id_source), external_id=toupper(external_id)) %>%
  unique()
glimpse(drugbank_drug_external_df)
```

### Map files

Map OTP indications to DrugBank drug targets. Filter by drugs that are maximum in phase 3 or 4, to make sure they are effective drugs.

```{r}
# Merge
otp_indications_to_drugbank_targets = otp_indications_df %>% 
  inner_join(drugbank_drug_external_df, by=c("chembl_id"="external_id")) %>% 
  inner_join(drugbank_drug_targets_df, by=c("DB_ID"="DB_ID")) %>%
  filter(max_phase_indication >= 3) %>% 
  unique()
head(otp_indications_to_drugbank_targets)
# Save
otp_indications_to_drugbank_targets_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/drug_targets/otp_indications_to_drugbank_targets.txt"
otp_indications_to_drugbank_targets %>% fwrite(otp_indications_to_drugbank_targets_file)
```

### Create drug target files for specific diseases

Breast cancer:

```{r}
unique(otp_indications_to_drugbank_targets$efo_term[grepl("breast", otp_indications_to_drugbank_targets$efo_term)])
```

```{r}
breast_cancer_drug_targets = otp_indications_to_drugbank_targets %>%
  dplyr::filter(efo_term == "breast neoplasm") %>%
  dplyr::select(DB_Name, Gene_Target) %>%
  dplyr::rename("ID"="DB_Name", "Target"="Gene_Target") %>%
  unique()
head(breast_cancer_drug_targets)
length(unique(breast_cancer_drug_targets$ID))
length(unique(breast_cancer_drug_targets$Target))
nrow((breast_cancer_drug_targets %>% dplyr::select(ID, Target) %>% unique()))
# Save
breast_cancer_drug_targets_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/drug_targets/drugbank_targets_breast.neoplasm.txt"
breast_cancer_drug_targets %>% fwrite(breast_cancer_drug_targets_file)
```

Breast cancer:

```{r}
unique(otp_indications_to_drugbank_targets$efo_term[grepl("bowel", otp_indications_to_drugbank_targets$efo_term)])
```

```{r}
# Using indications from OTP
ibd_drug_targets_otp = otp_indications_to_drugbank_targets %>%
  dplyr::filter(efo_term == "inflammatory bowel disease") %>%
  dplyr::select(DB_Name, Gene_Target) %>%
  dplyr::rename("ID"="DB_Name", "Target"="Gene_Target") %>%
  unique()
head(ibd_drug_targets_otp)
length(unique(ibd_drug_targets_otp$ID))
length(unique(ibd_drug_targets_otp$Target))
nrow((ibd_drug_targets_otp %>% dplyr::select(ID, Target) %>% unique()))
# Using indications from DrugBank
drugbank_drug_targets_df %>% dplyr::filter(grepl("inflammatory bowel disease", Indication)) %>% select(DB_Name, Indication) %>% unique()
ibd_drug_targets_db = drugbank_drug_targets_df %>%
  dplyr::filter(grepl("inflammatory bowel disease", Indication)) %>%
  dplyr::select(DB_Name, Gene_Target) %>%
  dplyr::rename("ID"="DB_Name", "Target"="Gene_Target") %>%
  unique()
head(ibd_drug_targets_db)
length(unique(ibd_drug_targets_db$ID))
length(unique(ibd_drug_targets_db$Target))
nrow((ibd_drug_targets_db %>% dplyr::select(ID, Target) %>% unique()))
# Merge both methods together
ibd_drug_targets = rbind(ibd_drug_targets_otp, ibd_drug_targets_db) %>% unique()
head(ibd_drug_targets)
length(unique(ibd_drug_targets$ID))
length(unique(ibd_drug_targets$Target))
nrow((ibd_drug_targets %>% dplyr::select(ID, Target) %>% unique()))
# Save
ibd_drug_targets_file = "/home/j.aguirreplans/Projects/Scipher/SampleSize/data/drug_targets/drugbank_targets_inflammatory.bowel.disease.txt"
ibd_drug_targets %>% fwrite(ibd_drug_targets_file)
```

