---
title: "GTEx preprocessing"
author: "Joaquim Aguirre-Plans"
date: "26/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Preprocess GTEx data (version v8).

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
set.seed(1510)
```

## Read files

```{r define_files}
# Define input files
samples_subjects_file = '/home/j.aguirreplans/Databases/GTEx/v8/GTEx_Annotations_SamplesSubjectsMerged.txt'
tpm_file = '/home/j.aguirreplans/Databases/GTEx/v8/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_tpm.gct'

# Define output file
tpm_filt_file = '/home/j.aguirreplans/Databases/GTEx/v8/GTEx_RNASeq_gene_tpm_filtered.gct'
```

```{r read_files}
# Read input files
samples_df = fread(samples_subjects_file)
tpm_df = fread(tpm_file)

```

## Remove genes with low TPM expression

We filter the samples in the TPM file by keeping the ones that are in the samples information file:

```{r filter_tpm_file_samples}
dim(tpm_df[,-1:-2])
tpm_df <- tpm_df %>% select(c("Name", "Description", samples_df$SAMPID))
dim(tpm_df[,-1:-2])
length(samples_df$SAMPID)

```

We calculate the median, mean, sum and standard deviation of the expression of each gene:

```{r}
info_df = data.frame(Name=tpm_df$Name, Description=tpm_df$Description, sum=rowSums(tpm_df[,-1:-2]))
info_df$sd = apply(tpm_df[,-1:-2], 1, sd)
info_df$mean = rowMeans(tpm_df[,-1:-2])
info_df$median = apply(tpm_df[,-1:-2], 1, median)
info_df$log2_median = log2(info_df$median)
info_df$log2_mean = log2(info_df$mean)
info_df$zero_sum_counts = as.numeric(apply(tpm_df[,-1:-2], 1, function(i) sum(i == 0) )) # Number of times that the sum of the TPMs is equal to 0

dim(info_df)
```

We keep the genes whose sum, standard deviation, mean or median expression is above 0.
Also, we keep the genes whose expression is 0 in more than 50% of the samples (as in https://www.biostars.org/p/377059/):

```{r}
info_filt_df = info_df[(info_df$sum > 0) & (info_df$mean > 0) & (info_df$median > 0) & (info_df$sd > 0) & (info_df$zero_sum_counts < dim(tpm_df[, -1])[2]/2) ,]
dim(info_filt_df)

```

We filter the TPM matrix:

```{r}
tpm_filt_df = tpm_df[tpm_df$Name %in% info_filt_df$Name,]
dim(tpm_filt_df)

```

Save the filtered gene expression data:

```{r}
fwrite(tpm_filt_df, tpm_filt_file)

```

```{r}
rm(tpm_df)
rm(tpm_filt_df)
rm(info_df)
rm(info_filt_df)

```

