---
title: "PPI preprocessing"
author: "Joaquim Aguirre-Plans"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Create a version of the PPI network that is ready to be used in my analyses.

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(igraph)
library(wTO)
set.seed(1510)
```

### Read files

```{r define_files}
# Define working directories
ppi_dir = '/home/j.aguirreplans/data/PPI'

# Define input files
ppi_network_file = paste(ppi_dir, 'interactome_2019_merged.csv', sep='/')
ppi_info_file = paste(ppi_dir, 'interactome_2019_merged_protAnnots.csv', sep='/')

# Define output file
ppi_network_symbol_file = paste(ppi_dir, 'interactome_2019_merged_symbols.csv', sep='/')

```


### Merge gene IDs and symbols

```{r map_symbols_to_ppi}
PPI = fread(ppi_network_file)
PPI_info_df = fread(ppi_info_file)
head(PPI)
head(PPI_info_df)
PPI_symbol = left_join(PPI[,c("proteinA", "proteinB")], PPI_info_df[,c("GeneID", "Symbol")], by=c("proteinA"="GeneID")) %>% rename(proteinA_symbol=Symbol)
PPI_symbol = left_join(PPI_symbol, PPI_info_df[,c("GeneID", "Symbol")], by=c("proteinB"="GeneID")) %>% rename(proteinB_symbol=Symbol)
head(PPI_symbol)

```

The initial PPI network has `r nrow(PPI_symbol)` edges and `r length(unique(c(PPI_symbol$proteinA_symbol, PPI_symbol$proteinB_symbol)))` nodes.


### Remove self-loops, repeated edges and isolated nodes

```{r select_unique_links}
PPI_symbol = PPI_symbol %>% unique()
dim(PPI_symbol)
PPI_symbol = PPI_symbol %>%
  filter((proteinA_symbol != "") & (proteinB_symbol != "")) %>%
  filter(proteinA_symbol != proteinB_symbol)
dim(PPI_symbol)

```

The PPI network, after removing self-loops, repeated edges and isolated nodes, has `r nrow(PPI_symbol)` edges and `r length(unique(c(PPI_symbol$proteinA_symbol, PPI_symbol$proteinB_symbol)))` nodes.


### Select the Largest Connected Component (LCC)

```{r select_lcc}
# Get LCC
ppi_net = graph_from_data_frame((PPI_symbol %>% dplyr::select(proteinA_symbol, proteinB_symbol)), directed=F) %>% simplify()
print(gsize(ppi_net))
print(gorder(ppi_net))
components_net = components(ppi_net)
ppi_net = induced.subgraph(ppi_net, vids = V(ppi_net)[components_net$membership == which.max(components_net$csize)] )
print(gsize(ppi_net))
print(gorder(ppi_net))
lcc_nodes = V(ppi_net)$name

# Filter interactions not in LCC
PPI_symbol = PPI_symbol %>% filter((proteinA_symbol %in% lcc_nodes) & (proteinB_symbol %in% lcc_nodes))

```

The PPI network, after keeping only the interactions in the LCC, has `r nrow(PPI_symbol)` edges and `r length(unique(c(PPI_symbol$proteinA_symbol, PPI_symbol$proteinB_symbol)))` nodes.


### Order the names in the edges in alphabetical order

```{r order_names_in_edges}
# Get a matrix of distances between PPIs
ppi_distances = distances(ppi_net)
ppi_distances[1:10,1:10]

# Order the rows and columns of the matrix alphabetically
ppi_distances = ppi_distances[order(rownames(ppi_distances)), order(colnames(ppi_distances))]
ppi_distances[1:10,1:10]

# Convert the matrix to an in-line format and order them by name
ppi_distances = ppi_distances %>% wTO.in.line() %>% rename(distances=wTO)
head(ppi_distances)

# Select only the direct interactions
PPI = ppi_distances %>% filter(distances == 1) %>% select(Node.1, Node.2) %>% rename("proteinA_symbol" = "Node.1", "proteinB_symbol" = "Node.2")
rm(ppi_distances)
head(PPI)

PPI_ordered = left_join(PPI, PPI_info_df[,c("GeneID", "Symbol")], by=c("proteinA_symbol"="Symbol")) %>% rename(proteinA=GeneID)
PPI_ordered = left_join(PPI_ordered, PPI_info_df[,c("GeneID", "Symbol")], by=c("proteinB_symbol"="Symbol")) %>% rename(proteinB=GeneID)
PPI_ordered = PPI_ordered[,c("proteinA", "proteinB", "proteinA_symbol", "proteinB_symbol")]
head(PPI_ordered)

# Safe the final file
fwrite(PPI_ordered, ppi_network_symbol_file)

```

The PPI network, after ordering the names of the edges, has `r nrow(PPI_ordered)` edges and `r length(unique(c(PPI_ordered$proteinA_symbol, PPI_ordered$proteinB_symbol)))` nodes.
