---
title: "Scipher subsampling"
author: "Joaquim Aguirre-Plans"
date: "1/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Create subsamples of gene expression data from GTEx, with the aim of testing the effect of sample size in these subsamples.

```{r, message=FALSE}
library(data.table)
library(dplyr)
set.seed(1510)
```

## Read files

```{r define_files}
# Define working directories
databases_dir = '/home/j.aguirreplans/Databases'
ppi_dir = '/home/j.aguirreplans/data/PPI'
data_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data/Dec2021'
output_dir = '/home/j.aguirreplans/Projects/Scipher/SampleSize/data'

# Define input files and dirs
metadata_RNA_file = paste(data_dir, '00_data/scipher_metadata_rnaseq_counts.csv', sep='/')
metadata_RNA_unique_patients_file = paste(data_dir, '00_data/scipher_unique_patients_metadata_rnaseq_counts.csv', sep='/')
metadata_RNA_unique_patients_0m_file = paste(data_dir, '00_data/scipher_unique_patients_metadata_rnaseq_counts_0m.csv', sep='/')

```

```{r read_files}
# Read input files
metadata_RNA = fread(metadata_RNA_file)
metadata_RNA_unique_patients = fread(metadata_RNA_unique_patients_file)
metadata_RNA_unique_patients_0m = fread(metadata_RNA_unique_patients_0m_file)

```

## Subsampling

```{r sampling_function}
## Random sampling with replacement
create_random_subsamples_with_replacement <- function(samples_list, size_list, rep, output_dir, output_file_name){
  subsamples_df = data.frame(matrix(ncol=3,nrow=0, dimnames=list(NULL, c("size", "rep", "samples"))))
  seed <- 1
  for(size in size_list){
    set.seed(seed)
    seed <- seed + 1
    for(r in 1:rep){
      subsample <- sort(sample(samples_list, size=size, replace=TRUE))
      subsample_df <- data.frame(samples=subsample)
      output_file = paste(output_dir, paste(paste(output_file_name, 'size', size, 'rep', r, sep='_'), '.txt', sep=''), sep='/')
      subsample_df %>% write.table(output_file, quote = F, sep='\t', row.names = F, col.names = T)
      subsamples_df = rbind(subsamples_df, data.frame(size=size, rep=r, samples=paste(subsample, collapse = ';')))
    }
  }
  return(subsamples_df)
}

```

> Important to note: There are many samples that have the same Scipher_id, meaning that they come from the same patient. They only differ by the "batch", which can be either "seq_fac1" or "seq_fac2".


```{r}
# Select all samples
complete_dataset = metadata_RNA$Scipher_id.visit.batch

# Select all samples of unique patients at different visits
sample_per_patient_all_visits = metadata_RNA_unique_patients$Scipher_id.visit

# Select all samples of unique patients at baseline
sample_per_patient_baseline = metadata_RNA_unique_patients_0m$Scipher_id.visit

# Get the responder and nonresponder samples at baseline
sample_per_patient_baseline_df = metadata_RNA_unique_patients_0m %>% filter(Scipher_id.visit %in% sample_per_patient_baseline)

# Responders are selected when they respond in eular and do not have any non-response
responder_euler_df = sample_per_patient_baseline_df %>% filter(( (eular_binary_3m == "responder") & (eular_binary_6m == "responder")) | ((eular_binary_3m == "responder") & (eular_binary_6m == "")) | ((eular_binary_3m == "") & (eular_binary_6m == "responder")) ) # Response in eular and not any non-response
#responder_df = sample_per_patient_baseline_df %>% filter( (!(eular_binary_3m == "nonresponder")) & (!(eular_binary_6m == "nonresponder")) & (!(acr50_3m == "nonresponder")) & (!(acr50_6m == "nonresponder")) & ( (eular_binary_3m == "responder") | (eular_binary_6m == "responder") | (acr50_3m == "responder") | (acr50_6m == "responder") ) ) # Response either in eular or acr50 and not any non-response
sample_per_responder_baseline = responder_euler_df$Scipher_id.visit

# Nonresponders are selected when they don't respond in eular and do not have any response
nonresponder_euler_df = sample_per_patient_baseline_df %>% filter(( (eular_binary_3m == "nonresponder") & (eular_binary_6m == "nonresponder")) | ((eular_binary_3m == "nonresponder") & (eular_binary_6m == "")) | ((eular_binary_3m == "") & (eular_binary_6m == "nonresponder")) )
#nonresponder_df = sample_per_patient_baseline_df %>% filter( (!(eular_binary_3m == "responder")) & (!(eular_binary_6m == "responder")) & (!(acr50_3m == "responder")) & (!(acr50_6m == "responder")) & ( (eular_binary_3m == "nonresponder") | (eular_binary_6m == "nonresponder") | (acr50_3m == "nonresponder") | (acr50_6m == "nonresponder") ) ) # Non-response in eular and not any response
sample_per_nonresponder_baseline = nonresponder_euler_df$Scipher_id.visit

```

We have selected different groups of samples according to different criteria:

* All samples, no matter the visit or batch (`r length(complete_dataset)`)
* One sample per patient at different visits (`r length(sample_per_patient_all_visits)`)
* One sample per patient at baseline (`r length(sample_per_patient_baseline)`)
* One sample per responder patient at baseline (`r length(sample_per_responder_baseline)`)
* One sample per non-responder patient at baseline (`r length(sample_per_nonresponder_baseline)`)


```{r}
group_to_samples = list("complete_dataset"=complete_dataset, "sample_per_patient_all_visits"=sample_per_patient_all_visits, "sample_per_patient_baseline"=sample_per_patient_baseline, "responder_baseline"=sample_per_responder_baseline, "nonresponder_baseline"=sample_per_nonresponder_baseline)
n_repetitions = 10

for (group in names(group_to_samples)){
  sampling_dir = paste(output_dir, 'sampling/Scipher/Dec2021/sampling_with_repetition', group, sep='/')
  dir.create(sampling_dir, showWarnings = FALSE)
  samples = group_to_samples[[group]]
  size_list <- seq(10, length(samples), 10)
  # Create a file with a list of all samples
  all_samples_df <- data.frame(samples=samples)
  output_file = paste(sampling_dir, paste('scipher_', group, '_all_samples.txt', sep=""), sep='/')
  all_samples_df %>% write.table(output_file, quote = F, sep='\t', row.names = F, col.names = T)
  # Create files with lists of subsamples
  output_file_name = paste('scipher_', group, sep = "")
  subsamples_df <- create_random_subsamples_with_replacement(samples_list=samples, size_list=size_list, rep=n_repetitions, output_dir=sampling_dir, output_file_name=output_file_name)

}
```

