---
title: "TCGA analysis"
author: "Joaquim Aguirre-Plans"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Description

Analyze TCGA dataset from GDC.

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(forcats)
library(ggplot2)
require(magrittr)
library(tidyr)
library(igraph)
library(rjson)
set.seed(1510)
`%ni%` <- Negate(`%in%`)
```


```{r}
# Define working directories
gdc_dir = '/home/j.aguirreplans/Databases/TCGA/2022-03-28-Dataset/TCGA'
input_dir = paste(gdc_dir, 'raw', sep='/')
data_dir = paste(input_dir, 'data', sep='/')
metadata_dir = paste(input_dir, 'metadata', sep='/')
additional_dir = paste(input_dir, 'additional', sep='/')
output_dir = paste(gdc_dir, 'out', sep='/')
```

## Mappings

We read the mappings file:

```{r}
# Define and read mappings file
mappings_file = paste(additional_dir, 'files.2022-03-28.json', sep='/')
mappings_json <- fromJSON(paste(readLines(mappings_file), collapse=""))
```

```{r}
# Parse file and case IDs from mappings file
mappings_df = data.frame(matrix(ncol=2,nrow=0, dimnames=list(NULL, c("file_id", "case_id"))))
for(n in 1:length(mappings_json)){
    file_id = mappings_json[[n]]$file_name
    case_id = mappings_json[[n]]$cases[[1]]$case_id # All cases are composed of only 1 case id
    mappings_df <- rbind(mappings_df, data.frame(file_id=file_id, case_id=case_id))
}
```

```{r}
head(mappings_df)
```

```{r}
glimpse(mappings_df)
```

There are `r length(unique(mappings_df$file_id))` unique file IDs and `length(unique(mappings_df$case_id))` unique case IDs, meaning that there are case IDs that are mapped to multiple files.

```{r}
length(unique(mappings_df$file_id))
length(unique(mappings_df$case_id))
```

```{r}
mappings_txt_file = paste(output_dir, 'files.2022-03-28.txt', sep='/')
fwrite(mappings_df, mappings_txt_file, sep='\t')
```

```{r}
# Get duplicated case IDs
mappings_duplicated_df = mappings_df %>% filter(case_id %in% subset(mappings_df,duplicated(case_id))$case_id)
length(unique(mappings_duplicated_df$case_id))
length(unique(mappings_duplicated_df$file_id))
```

There are `r length(unique(mappings_duplicated_df$case_id))` case IDs that are duplicated in `length(unique(mappings_duplicated_df$file_id))` file IDs.


## Metadata

There are two folders of metadata in GDC:
- Clinical
    - `clinical.tsv`: Age, if there is treatment, and which type of treatment.
    - `exposure.tsv`: Exposure to substances such as alcohol or tobacco.
    - `family_history.tsv`: Empty.
- Biospecimen: Information about the samples and how they were extracted.
    - `aliquot.tsv`: Information about the aliquot
    - `analyte.tsv`: Information about the analyte.
    - `portion.tsv`: ??
    - `sample.tsv`: Information about the sample.
    - `slide.tsv`: Information about microscopic slide.

Let's analyze the `clinical.tsv` file in the **Clinical** folder:

```{r}
# Define clinical metadata files
clinical_dir = paste(metadata_dir, 'clinical', sep='/')
clinical_file = paste(clinical_dir, 'clinical.tsv', sep='/')
exposure_file = paste(clinical_dir, 'exposure.tsv', sep='/')
family_history_file = paste(clinical_dir, 'family_history.tsv', sep='/')

# Read clinical metadata files
clinical_df = fread(clinical_file, header=TRUE)
exposure_df = fread(exposure_file, header=TRUE)
family_history_df = fread(family_history_file, header=TRUE)
```

`clinical.tsv` contains `r ncol(clinical_df)` columns of different clinical information about `r length(unique(clinical_df$case_id))` case IDs in `r nrow(clinical_df)` rows:

```{r}
glimpse(clinical_df)
```

The most informative columns for the purpose of the project are:

- `case_id`
- `case_submitter_id`
- `project_id`
- `age_at_index`
- `ethnicity`
- `gender`
- `race`
- `vital_status`
- `year_of_birth`
- `age_at_diagnosis`
- `ajcc_clinical_m`
- `ajcc_clinical_n`
- `ajcc_clinical_stage`
- `ajcc_clinical_t`
- `ajcc_pathologic_m`
- `ajcc_pathologic_n`
- `ajcc_pathologic_stage`
- `ajcc_pathologic_t`
- `icd_10_code`
- `primary_diagnosis`
- `year_of_diagnosis`
- `treatment_or_therapy`
- `treatment_type`

Let's analyze the `sample.tsv` file in the **Biospecimen** folder:

```{r}
# Define clinical metadata files
biospecimen_dir = paste(metadata_dir, 'biospecimen', sep='/')
aliquot_file = paste(biospecimen_dir, 'aliquot.tsv', sep='/')
analyte_file = paste(biospecimen_dir, 'analyte.tsv', sep='/')
portion_file = paste(biospecimen_dir, 'portion.tsv', sep='/')
sample_file = paste(biospecimen_dir, 'sample.tsv', sep='/')
slide_file = paste(biospecimen_dir, 'slide.tsv', sep='/')

# Read clinical metadata files
aliquot_df = fread(aliquot_file, header=TRUE)
analyte_df = fread(analyte_file, header=TRUE)
portion_df = fread(portion_file, header=TRUE)
sample_df = fread(sample_file, header=TRUE)
slide_df = fread(slide_file, header=TRUE)
```

```{r}
glimpse(sample_df)
```

Let's merge the clinical data with the mappings:

```{r}
# Merge mappings with clinical
metadata_df = inner_join(mappings_df, clinical_df, by="case_id")
# Remove redundant columns
metadata_df = metadata_df[vapply(metadata_df, function(x) length(unique(x)) > 1, logical(1L))]
dim(metadata_df)
```

The merged metadata contains `r ncol(metadata_df)` columns of different clinical information about `r length(unique(metadata_df$case_id))` case IDs and `r length(unique(metadata_df$file_id))` file IDs in `r nrow(metadata_df)` rows:

```{r}
# Write the output file
metadata_file = paste(output_dir, 'metadata.txt', sep='/')
fwrite(metadata_df, metadata_file, sep='\t')
```

TCGA contains `r length(unique(metadata_df$project_id))` project IDs. Each project is focused on a different type of cancer. Each project contains the following number of case IDs:

```{r}
# Remove file IDs and keep only case IDs
metadata_cases_df = metadata_df %>% select(!file_id) %>% unique()

# Get table of project IDs
table(metadata_cases_df$project_id)
```

```{r}
# Plot project IDs
metadata_cases_count <- metadata_cases_df %>% select(case_id, project_id) %>% unique() %>% count(project_id)
metadata_cases_plot <- metadata_cases_count %>%
  mutate(project_id = fct_reorder(project_id, n)) %>%
  ggplot( aes(x=project_id, y=n)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()
#ggsave('/home/j.aguirreplans/Projects/Scipher/SampleSize/data/out/tcga_analysis/num_cases_per_project.png', plot=metadata_cases_plot, type='cairo', dpi = 300, height=8, width = 10)
metadata_cases_plot

```

For more information about each project, we can parse the projects file, which contains information about the primary sites and disease types of each project:

```{r}
projects_file = paste(metadata_dir, 'projects/projects.2022-03-29.json', sep='/')
projects_json <- fromJSON(paste(readLines(projects_file), collapse=""))

# Parse project names
projects_df = data.frame(matrix(ncol=3,nrow=0, dimnames=list(NULL, c("project_id", "primary_site", "disease_type"))))
for(n in 1:length(projects_json)){
    project_id = projects_json[[n]]$project_id
    primary_sites = projects_json[[n]]$primary_site
    disease_types = projects_json[[n]]$disease_type
    projects_df <- rbind(projects_df, data.frame(project_id=project_id, primary_site=paste(primary_sites, collapse="|"), disease_type=paste(disease_types, collapse="|")))
}

# Write the output file
projects_txt_file = paste(output_dir, 'projects.txt', sep='/')
fwrite(projects_df, projects_txt_file, sep='\t')

```

```{r}
glimpse(projects_df)
```


## Data

We read the processed file:

```{r}
# Define and read data file
tcga_file = paste(output_dir, 'TCGA.csv', sep='/')
tcga_df = fread(tcga_file, header=TRUE)
```

The `tcga_df` dataframe contains `r ncol(tcga_df)` columns (samples) and `r nrow(tcga_df)` rows (genes).

```{r}
tcga_df %<>% rename("Ensembl"="V1")
head(tcga_df[1:10,1:5])
```

```{r}
dim(tcga_df)
```

The columns (samples) are actually file IDs. We select the ones present in the metadata:

```{r}
metadata_file_ids = unique(metadata_df$file_id)
tcga_df = cbind(tcga_df$Ensembl, tcga_df %>% select(all_of(metadata_file_ids)))
tcga_df %<>% rename("Ensembl"="V1")
dim(tcga_df)
```

Now it contains `r ncol(tcga_df)` columns (samples) and `r nrow(tcga_df)` rows (genes).


### Gene pre-processing

In the gene expression dataframe, there are `r nrow(tcga_df)` genes in Ensembl notation. We will transform these genes into HGNC gene symbol notation.

First we read the file `multiple_hgnc_dict_set_2022-01-01.txt`, containing gene symbols and their aliases in other notations:

```{r}
gene_annotations_dir = '/home/j.aguirreplans/data/gene_annotations/data/out'
gene_dict_file = paste(gene_annotations_dir, 'multiple_hgnc_dict_set_2022-01-01.txt', sep='/')
gene_dict_df = fread(gene_dict_file) %>% unique()
gene_dict_df = gene_dict_df[!((gene_dict_df$HGNC_Symbol=="")|(gene_dict_df$alias=="")),]
gene_dict_df$HGNC_Symbol = toupper(gene_dict_df$HGNC_Symbol)
gene_dict_df$alias = toupper(gene_dict_df$alias)
gene_dict_df %<>% unique()
head(gene_dict_df)
```

We remove the Ensembl ID part after the "." in the gene expression dataframe:

```{r}
tcga_df$Ensembl = sub("\\..*", "", tcga_df$Ensembl)
head(tcga_df[1:5,1:5])
```

We use this dataframe to map Ensembl IDs to HGNC symbol IDs:

```{r}
tcga_df = inner_join(tcga_df, gene_dict_df, by=c("Ensembl"="alias"))
tcga_df = cbind((tcga_df %>% select("Ensembl", "HGNC_Symbol")), (tcga_df %>% select(-"Ensembl", -"HGNC_Symbol"))) %>% unique()
head(tcga_df[1:5,1:5])
```

```{r}
dim(tcga_df)
```

Now, we have `r length(unique(tcga_df$HGNC_Symbol))` unique HGNC symbols mapped to `r length(unique(tcga_df$Ensembl))` unique Ensembl IDs from a total of `r nrow(tcga_df)` rows. The problem is that there are Ensembl IDs mapped to the same HGNC code and viceversa. So, we have to do the following procedures:
- If an Ensembl ID has multiple HGNC symbols, keep only one symbol
- If a HGNC symbol has multiple Ensembl IDs, calculate the median

We get the HGNC symbols with multiple Ensembl IDs:

```{r}
repeated_ensembl = unique(tcga_df$Ensembl[duplicated(tcga_df$Ensembl)])
length(repeated_ensembl)
repeated_ensembl_df = tcga_df[tcga_df$Ensembl %in% repeated_ensembl,][,c("Ensembl", "HGNC_Symbol")]
repeated_ensembl_df$HGNC_length = nchar(repeated_ensembl_df$HGNC_Symbol)
repeated_ensembl_df = repeated_ensembl_df[order(Ensembl, HGNC_length),]
head(repeated_ensembl_df)
```

We keep the smallest symbol, and if not, the first (in alphabetical order):

```{r}
repeated_ensembl_distinct_df = repeated_ensembl_df %>% distinct(Ensembl, .keep_all = TRUE) %>% select(Ensembl, HGNC_Symbol)
head(repeated_ensembl_distinct_df)
hgnc_to_remove = repeated_ensembl_df$HGNC_Symbol[!(repeated_ensembl_df$HGNC_Symbol %in% repeated_ensembl_distinct_df$HGNC_Symbol)]
tcga_df %<>% filter(!(HGNC_Symbol %in% hgnc_to_remove))
```

```{r}
dim(tcga_df)
```

We get the Ensembl IDs with multiple HGNC symbols:

```{r}
repeated_hgnc = unique(tcga_df$HGNC_Symbol[duplicated(tcga_df$HGNC_Symbol)])
length(repeated_hgnc)
repeated_hgnc_df = tcga_df[tcga_df$HGNC_Symbol %in% repeated_hgnc,][,c("Ensembl", "HGNC_Symbol")][order(HGNC_Symbol),]
head(repeated_hgnc_df)
```

We calculate the median for the HGNC genes with multiple Ensembl IDs:

```{r}
tcga_df = subset(tcga_df, select=-Ensembl)
repeated_hgnc_aggregated = aggregate(.~HGNC_Symbol, tcga_df[tcga_df$HGNC_Symbol %in% repeated_hgnc,], median)
tcga_df = rbind(tcga_df[!(tcga_df$HGNC_Symbol %in% repeated_hgnc),], repeated_hgnc_aggregated)
dim(tcga_df)
```

### Sample pre-processing

There are case IDs that are mapped to multiple files. We will merge the case IDs that have multiple file IDs, and substitute the file IDs with case IDs in the data file.

We identify the case IDs with multiple file IDs:

```{r}
mappings_tcga_df = mappings_df %>% filter(file_id %in% colnames(tcga_df)[-1])
mappings_tcga_duplicated_df = mappings_tcga_df %>% filter(case_id %in% subset(mappings_tcga_df,duplicated(case_id))$case_id)
head(mappings_tcga_duplicated_df)
```

There are `r length(unique(mappings_tcga_duplicated_df$case_id))` case IDs with multiple file IDs. We calculate the median between the expression of these file IDs.

```{r}
genes_tcga_df = tcga_df$HGNC_Symbol
tcga_df %<>% select(-HGNC_Symbol) %>% t() %>% as.data.frame()
colnames(tcga_df) = genes_tcga_df
tcga_df = cbind(data.frame(file_id=row.names(tcga_df)), tcga_df)
tcga_df %<>% inner_join(mappings_df, by=c("file_id"="file_id"))
tcga_df %<>% relocate(file_id, case_id)
tcga_df = subset(tcga_df, select=-file_id)
case_ids_aggregated = aggregate(.~case_id, tcga_df[tcga_df$case_id %in% unique(mappings_tcga_duplicated_df$case_id),], median)
tcga_df = rbind(tcga_df[!(tcga_df$case_id %in% unique(mappings_tcga_duplicated_df$case_id)),], case_ids_aggregated)
dim(tcga_df)

```

```{r}
# We transpose the matrix
genes = colnames((tcga_df %>% select(-case_id)))
case_ids = tcga_df$case_id
tcga_df = t(as.matrix((tcga_df %>% select(-case_id)))) %>% as.data.frame()
colnames(tcga_df) = case_ids
tcga_df = cbind(data.frame(Gene=genes), tcga_df)
```

```{r}
tcga_processed_file = paste(output_dir, 'TCGA_processed.csv', sep='/')
fwrite(tcga_df, tcga_processed_file, sep='\t')
```

The processed `tcga_df` dataframe contains `r as.integer(ncol(tcga_df)-1)` columns (case IDs) and `r nrow(tcga_df)` rows (genes).

```{r}
head(tcga_df[1:10,1:5])
```

