---
title: "RA - Generate Data for wTO"
output:
  html_document:
    theme: united
    highlight: tang
author: Deisy Morselli Gysi
---
#### Northeastern University
#### Project: Scipher RA - Responder vs Non-responders to TNF drugs

## Description
Co-expression Networks have to be build separately for each condition before their are compared. For that, we need to create four networks. 

The data has to be split in 4 data sets for the co-expression networks:

 - Responders: 3 and 6 m
 - Non responders: 3 and 6 m

```{r, message=FALSE}
require(dplyr)
require(magrittr)
require(data.table)
```


```{r}
metadata_RNA = fread("~/Desktop/PostDoc/00_Projects/RA/00_Data/0911/Processed/Meta_RNA_Cleaned.csv")

Responder_3m = metadata_RNA %>% filter(acr50_3m == "responder")
Responder_6m = metadata_RNA %>% filter(acr50_6m == "responder")

NResponder_3m = metadata_RNA %>% filter(acr50_3m == "nonresponder")
NResponder_6m = metadata_RNA %>% filter(acr50_6m == "nonresponder")

R3 = Responder_3m[, -c(1:52)] %>% t %>% as.data.frame() 
R6 = Responder_6m[, -c(1:52)] %>% t %>% as.data.frame() 
 
nR3 = NResponder_3m[, -c(1:52)] %>% t %>% as.data.frame() 
nR6 = NResponder_6m[, -c(1:52)] %>% t%>% as.data.frame() 
```


```{r}
R3 %>% fwrite(., "../data/out/Responder_3m.csv", row.names = TRUE)
R6 %>%  fwrite(., "../data/out/Responder_6m.csv", row.names = TRUE)

nR3 %>% fwrite(., "../data/out/NonResponder_3m.csv", row.names = TRUE)
nR6 %>%  fwrite(., "../data/out/NonResponder_6m.csv", row.names = TRUE)
```


Next run the following
```
# 
R CMD BATCH --vanilla '--args  N=1000 save="../data/out/wTO/Responder_3m.csv" file="../data/out/Responder_3m.csv"' ../code/99_Calls_wTO.R ../code/log/Responder_3m.out &
#   
R CMD BATCH --vanilla '--args  N=1000 save="../data/out/wTO/Responder_6m.csv" file="../data/out/Responder_6m.csv"' ../code/99_Calls_wTO.R ../code/log/Responder_6m.out &
#
R CMD BATCH --vanilla '--args  N=1000 save="../data/out/wTO/NonResponder_3m.csv" file="../data/out/NonResponder_3m.csv"' ../code/99_Calls_wTO.R ../code/log/NonResponder_3m.out &
#
R CMD BATCH --vanilla '--args  N=1000 save="../data/out/wTO/NonResponder_6m.csv" file="../data/out/NonResponder_6m.csv"' ../code/99_Calls_wTO.R ../code/log/NonResponder_6m.out &
```
## Session Info
```{r}
sessioninfo::session_info()
```