---
title: "Biomarkers Proximity"
output:
  html_document:
    theme: united
    highlight: tang
author: Deisy Morselli Gysi
---
#### Northeastern University
#### Project: Scipher RA - Responder vs Non-responders to TNF drugs

## Description
The aim here is to perform all the Network Medicine Analysis for the set of genes found. 
Not that, the enrichment for them are done at [Enrichr](https://maayanlab.cloud/Enrichr/). 


```{r, message=FALSE}
require(tidyr)
require(data.table)
require(magrittr)
require(dplyr)
require(methods)
require(magrittr)
require(tidyverse)
require(stringr)
require(ggplot2)
require(igraph)
require(progress)
require(NetSci)
`%ni%` <- Negate(`%in%`)
```

## Calculate the LCC for each biomarker group

```{r}
N = 1000  #Simulations

PPI = fread("../data/PPI_Symbol_Entrez.csv")

gPPI = PPI[, c(3,4)] %>%
  graph_from_data_frame(., directed = F) %>%
  simplify() %>%
  delete.vertices(v="")

Genes_nR = fread("../data/out/CoDINA/CN_nR.txt", header = F)
Genes_R = fread("../data/out/CoDINA/CN_R.txt", header = F)
Genes_RnR = fread("../data/out/CoDINA/CN_a.txt", header = F)
###
### add genes known to RA
Genes_RA = fread("../data/RA_Genes.txt", header = F)
Genes_RA %<>% filter(V1 %in% V(gPPI)$name)
```




```{r}
LCC_nR = LCC_Significance(N = N, 
                          Targets = Genes_nR$V1,
                          G = gPPI)
Histogram_LCC(LCC_nR)

LCC_R = LCC_Significance(N = N, 
                         Targets = Genes_R$V1,
                         G = gPPI)
Histogram_LCC(LCC_R)

LCC_RnR = LCC_Significance(N = N, 
                           Targets = Genes_RnR$V1,
                           G = gPPI)
Histogram_LCC(LCC_RnR)

LCC_RA = LCC_Significance(N = N, 
                          Targets = Genes_RA$V1,
                          G = gPPI)
Histogram_LCC(LCC_RA)
```

## Calculate the separation across the data 

```{r}
Genes_nR$type = "non resp"
Genes_R$type = "resp"
Genes_RnR$type = "common"
Genes_RA$type = "RA"

All = rbind(Genes_nR, Genes_R, Genes_RnR, Genes_RA)

Jac = Jaccard(All[,c(2,1)])
Jac

sab = separation(gPPI, All[,c(2,1)])
S = sab$Sab
S[lower.tri(S)] = t(S)[lower.tri(S)]
```


```{r}
S %>% superheat::superheat(., 
                           pretty.order.rows = TRUE, 
                           pretty.order.cols = TRUE, 
                           heat.col.scheme = "red")


```


## Calculate the proximity to Scipher selected drugs

```{r}
ScipherDrugs = readxl::read_excel("../data/DrugScipher.xlsx")

DB = fread("../data/DB_DrugTargets_1201.csv")

DB %<>% 
  filter(Gene_Target %in% V(gPPI)$name) %>%
  filter(Name %in% ScipherDrugs$Drug) %>%
  select(Name, Gene_Target) %>% 
  unique()

# Amount of targets per drug
DB %>%
  group_by(Name) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```
```{r}
# Calculate the proximity
# 
names(DB) = c('ID', "Target" )
```

```{r}
proximity_RA = avr_proximity_multiple_target_sets(
  set = unique(DB$ID),
  G = gPPI,
  ST = DB,
  source = Genes_RA$V1,
  N = N
)
proximity_RA

proximity_RA %>% 
  filter(p.adjust(p_lt) < 0.001)
```


```{r}
proximity_nR = avr_proximity_multiple_target_sets(
  set = unique(DB$ID),
  G = gPPI,
  ST = DB,
  source = Genes_nR$V1,
  N = N
)
proximity_nR

proximity_nR %>% 
  filter(p.adjust(p_lt) < 0.001)

```

```{r}
proximity_nR_RnR = avr_proximity_multiple_target_sets(
  set = unique(DB$ID),
  G = gPPI,
  ST = DB,
  source = c(Genes_nR$V1, Genes_RnR$V1),
  N = N
)
proximity_nR_RnR

proximity_nR_RnR %>% 
  filter(p.adjust(p_lt) < 0.001)

```
```{r}
proximity_R = avr_proximity_multiple_target_sets(
  set = unique(DB$ID),
  G = gPPI,
  ST = DB,
  source = Genes_R$V1,
  N = N
)
proximity_R

proximity_R %>% 
  filter(p.adjust(p_lt) < 0.001)

```

```{r}
proximity_R_RnR = avr_proximity_multiple_target_sets(
  set = unique(DB$ID),
  G = gPPI,
  ST = DB,
  source = c(Genes_R$V1, Genes_RnR$V1),
  N = N
)
proximity_R_RnR

proximity_R_RnR %>% 
  filter(p.adjust(p_lt) < 0.001)
```

## Session Info 

```{r}
sessioninfo::session_info()
```

